### 2023-02-25 summary.tex
Start on a summary of operads

\documentclass{amsart}

\title{Summary of the things that I learned}
\author{Arnoud van der Leer}

\begin{document}
  \maketitle

  \section{Week 08}
  \subsection{Univalent Categories}
  A univalent category is a category in which the univalence axiom holds. I.e., a category $ \mathcal C $ in which, for all $ A, B \in \mathcal C_0 $, the canonical map $ (A =_{\mathcal C} B) \to (A \cong B) $ is an equivalence.

  \subsection{Categories}
  An \textbf{$ n $-category} is a category with $ 0 $-cells (objects), $ 1 $-cells (morphisms), $ 2 $-cells (morphisms between morphisms), up to $ n $-cells and various compositions: $ A \to B \to C $. $ A \xrightarrow{f, g, h} B $, $ f \Rightarrow g \Rightarrow h $. $ A \xrightarrow{f, g} B \xrightarrow{f^\prime, g^\prime} C $, $ \alpha: f \Rightarrow g $, and identities $ \alpha^\prime: f^\prime \Rightarrow g^\prime $ gives $ \alpha^\prime * \alpha: f^\prime \circ f \Rightarrow g^\prime \circ g $. These all need to work together `nicely'. An $ \omega $-category is the same, but all the way up.

  A topological space gives a (weak) $ \omega $-category. $ 0 $-cells are points, $ 1 $-cells are paths, $ 2 $-cells are homotopies etc. Composition is by glueing. It is a `groupoid', in the sense that all homotopies of dimension $ \geq 1 $ are invertible. However, glueing is not associative, so it is a `weak' $ \omega $-category.

  A category with only one object $ \star $ is equivalent to a monoid (with elements being the set $ \mathcal C(\star, \star) $). A $ 2 $-category with only one $ 0 $-cell is the same thing as a monoidal category (objects: the $ 1 $-cells. Morphisms: the $ 2 $-cells). A monoidal category with just one object gives $ 2 $ monoid structures on its set of morphisms. These are the same, and commutative.

  A \textbf{monoid} is a set with a multiplication and a unit. A \textbf{monad} on a category $ \mathcal A $ is a functor $ \mathcal A \to \mathcal A $, together with natural transformations $ \mu : T \circ T \to T $ (satisfying associativity) and $ \eta : 1_{\mathcal A} \to T $ (acting as a two-sided unit).

  A \textbf{presheaf} on a category $ \mathcal A $ is a functor $ \mathcal A^{\mathop{opp}} \to \mathbf{Set} $.

  Given a category $ \mathcal E $ and an object $ E \in \mathcal E_0 $, the \textbf{slice category} $ \mathcal E / E $ with objects being the maps $ D \xrightarrow{p} E $ and morphisms being commutative triangles.

  A \textbf{multicategory}, not necessarily the same as an $ n $-category, is a category in which arrows go from multiple objects to one, instead of from one object to one. I.e. it is a category with a class $ C_0 $ of objects, for all $ n $, and all $ a, a_1, \dots, a_n \in C_0 $, a class $ C(a_1, \dots, a_n; a) $ of `morphisms', and a composition 
  \[ C(a_1, \dots, a_n; a) \times C(a_{1, 1}, \dots, a_{1, k_1}; a_1) \times \dots \times C(a_{n, 1}, \dots, a_{n, k_n}; a_n) \to C(a_{1, 1}, \dots, a_{n, k_n}; a), \]
  written $ (\theta, \theta_1, \dots, \theta_n) \mapsto \theta (\theta_1, \dots, \theta_n) $ and for each $ a \in C_0 $ an identity $ 1_a \in C(a; a) $. It must satisfy associativity
  \[ \theta \circ(\theta_1 \circ(\theta_{1, 1}, \dots, \theta_{1, k_1}), \dots, \theta_n \circ (\theta_{n, 1}, \dots, \theta_{n, k_n})) = (\theta \circ (\theta_1, \dots, \theta_n)) \circ (\theta_{1, 1}, \dots, \theta_{n, k_n}) \]
  and identity
  \[ \theta \circ (1_{a_1}, \dots, 1_{a_n}) = \theta = 1_a \circ \theta. \]

  A \textbf{map of multicategories} is a function $ f_0: C_0 \to C_0^\prime $ and maps $ C(a_1, \dots, a_n; a) \to C(f_0(a_1), \dots, f_0(a_n); f_0(a)) $, preserving composition and identities.

  For $ C $ a multicategory, a \textbf{$ C $-algebra} is a map from $ C $ into the multicategory $ \mathbf{Set} $ (with objects $ \mathrm{Set}_0 $ and maps $ \mathrm{Set}(a_1, \dots, a_n; a) = \mathrm{Set}(a_1 \times \dots \times a_n ; a) $). I.e., for each $ a \in C_0 $, a set $ X(a) $, and for each map $ \theta: a_1, \dots, a_n \to a $, a function $ X(\theta): X(a_1) \times X(a_n) \to X(a) $. An example is, for a multicategory $ C $, to take $ X(a) = C(; a) $ (maps from the empty sequence into $ a $).

  \subsection{Operads}
  \subsubsection{Definitions}
  An \textbf{operad} is a multicategory with only one object. More explicitly, an operad has a set $ P(k) $ for every $ k \in \mathbb N $, whose elements can be thought of as $ k $-ary operations. It also has, for all $ n, k_1, \dots, k_n \in \mathbb N $, a \textit{composition} function
  \[ P(n) \times P(k_1) \times \dots \times P(k_n) \to P(k_1 + \dots + k_n) \]
  and an element $ 1 = 1_P \in P(1) $ called the \textbf{identity}, satisfying
  \[ \theta \circ (1, 1, \dots, 1) = \theta = \theta \circ 1 \]
  for all $ \theta $, and
  \[ \theta \circ(\theta_1 \circ(\theta_{1, 1}, \dots, \theta_{1, k_1}), \dots, \theta_n \circ (\theta_{n, 1}, \dots, \theta_{n, k_n})) = (\theta \circ (\theta_1, \dots, \theta_n)) \circ (\theta_{1, 1}, \dots, \theta_{n, k_n}) \]
  for all $ \theta \in P(n) $, $ \theta_1 \in P(k_1) $, \dots, $ \theta_n \in P(k_n) $ and all $ \theta_{1, 1} \dots \theta_{n, k_n} $.

  A \textbf{morphism of operads} is a family
  \[ f_n : (P(n) \to Q(n))_{n \in \mathbb N} \]
  of functions, preserving composition and identities.

  A \textbf{$ P $-algebra} for $ P $ an operad, is a set $ X $ and, for each $ n $, and $ \theta \in P(n) $, a function $ \overline{\theta}: X^n \to X $, satisfying the evident axioms (identity is the identity function, the function of a composition is the composition of the functions?).

  \subsubsection{Examples}
  For any vector space $ V $, there is an operad with $ P(k) = V^{\otimes k} \to V $.

  The terminal operad $ 1 $ has $ P(n) = \{ \star_1 \} $ for all $ n $. An algebra for $ 1 $ is a set $ X $ together with a function $ X^n \to X $, denoted as $ (x_1, \dots, x_n) \mapsto (x_1 \cdot \dots \cdot x_n) $, satisfying
  \[ ((x_{1, 1} \cdot \dots x_{1, k_1}) \cdot \dots \cdot (x_{n, 1} \cdot \dots \cdot x_{n, k_n})) = (x_{1, 1} \cdot \dots \cdot x_{n, k_n}) \]
  and
  \[ x = (x). \]
  The category of $ 1 $-algebras is the category of monoids.

  There exist various sub-operads of $ 1 $. For example, the smallest one has $ P(1) = \{ \star \} $ and $ P(n) = \emptyset $ for $ n \not = 1 $.

  Or the operad with $ P(0) = \emptyset $ and $ P(n) = \{ \star_n \} $ for $ n > 0 $, which has semigroups as its algebras (sets with associative binary operations).

  The suboperad with $ P(n) = \{ \star_n \} $ exactly when $ n \leq 1 $ has as its algebras the pointed sets.

\end{document}