### 2023-10-19 report/planning.tex
Add a planning, add half of the proof of the equivalence between lambda algebras and lambda theories

\documentclass{amsart}

\usepackage{a4wide}

\title{Planning}

\begin{document}
  \maketitle

  \section{Global planning}
  \begin{tabular}{l|l|l|l}
    \textbf{Activity} & \textbf{Priority} & \textbf{Duration} & \textbf{Done by}\\\hline
    Prove isomorphism between theories and algebras & High & 2 weeks & November 3\\\hline
    Polish code and do write-up & High & 1 week & November 10\\\hline
    Research and write-up for category of retracts & Middle & 2 weeks & November 24\\\hline
    Formalization of category of retracts & Middle & 2 weeks & December 8\\\hline
    Polish code & Middle & 1 week & December 15\\\hline
    Formalization of isomorphism between $ U_A $ and $ \Lambda_A $ & Middle & 1 week & December 22\\\hline
    Write-up and polishing & Middle & 1 week & December 29\\\hline
    Christmas & High & 2 weeks & January 12\\\hline
    Unforeseen problems & High & 3 weeks & February 2
  \end{tabular}

  \section{Isomorphism}
  \begin{tabular}{r|l}
    1/2 day\phantom{s} & Initiality of $ \Lambda $\\
    1/2 day\phantom{s} & The functor from lambda theories to $ \Lambda $-algebras\\
    1/2 day\phantom{s} & Definition of the presheaf category\\
    1/2 day\phantom{s} & The object $ A(1) $\\
    1/2 day\phantom{s} & Exponentiability of $ A(1) $\\
    1/2 day\phantom{s} & $ A(1)^{A(1)} \cong A(2) $\\
    2 days & Isomorphism between $ L $ and $ E(A(1)) $\\
    1/2 day\phantom{s} & Faithful functor\\
    1/2 day\phantom{s} & Induced functor on presheaf categories\\
    1/2 day\phantom{s} & Preserves products\\
    1/2 day\phantom{s} & Map from $ E(A(1)) $ to $ E(A^\prime(1)) $\\
    1/2 day\phantom{s} & Map equals $ f $, so functor is full\\\hline
    \textbf{7 1/2 days}
  \end{tabular}

\end{document}