### 2023-11-01 talks/t04-dutch-cats/outline.tex
Update report, write talk for dutchcats

\documentclass{amsart}

\title{Semantics for the $ \lambda $-calculus}
\author{Arnoud van der Leer\\Delft University of Technology\\\texttt{arnoudvanderleer@gmail.com}}

\begin{document}
  \maketitle

  \section{Intro}

  Hi, it is so good to see you all here. For those who don't know me, I am Arnoud van der Leer, master's student at TU Delft, and am currently working on my master's thesis, supervised by Benedikt and Kobe.

  Today, I am going to talk about the paper `Classical lambda calculus in modern dress'. This is a paper from 2013, written by Cambridge professor Martin Hyland.
  In this paper, he talks about models for the $ \lambda $-calculus, and proves three `big' theorems about them.
  I will also talk a little bit about my thesis, which revolves around this paper. My job is to `annotate' this paper, so to speak.

  \tableofcontents

  So here is the outline of my talk. We are currently in the introduction.
  To prove things about the $ \lambda $-calculus and its models, we first need establish the way in which we will formally talk about the $ \lambda $-calculus. We will also talk a bit about denotational semantics.
  Then we will have a look at the three main theorems.
  The first main theorem is Scotts representation theorem, which shows that every model of the $ \lambda $-calculus can be obtained from some category.
  Hyland calls the second main theorem the fundamental theorem of the $ \lambda $-calculus. This is a somewhat stronger result, which says that there is an equivalence between models and denotations for the $ \lambda $-calculus.
  For me, the last main theorem seems to be the most cryptic one, to me. It states that if we have a model for the $ \lambda $-calculus, then create its `category of retracts', then this category is `relatively cartesian closed'. It basically says: if we have a model for the $ \lambda $-calculus, we can create a category in which we can do some form of dependent type theory.
  Lastly, I will talk a bit about my contribution to this, so annotating the paper, and especially mechanizing it.

  \section{Talking about the $\lambda$-calculus}

  So, first of all, in this talk, I will talk almost exclusively about the \textbf{untyped} $ \lambda $-calculus. The untyped $ \lambda $-calculus essentially is about collections that consist of only functions. With the objects of these collections, (so functions), we can do a couple of things. First of all, we can create new objects, which are just the `variables'. Then, given two of these objects, we can apply one to the other. Lastly, given a function with $ n + 1 $ free variables, we can create a function with $ n $ free variables by abstracting.

  Again, note that everything here is a function.

  We can impose two types of equalities on these terms: $ \beta $-equality and $ \eta $-equality. One is about what happens when you first abstract and then apply. The other is about what happens if you first apply to a variable and then abstract again.

  So a model for the $ \lambda $-calculus is something that exhibits this structure. Something thas has variables, application and abstraction, and maybe $ \beta $ and/or $ \eta $-equality.

  One such collection is the \textit{pure} $ \lambda $-calculus. One can view this as an inductive type, given by these three constructors. We can also create the pure $ \lambda $-calculus with $ \beta $-equality, which arises from this inductive type via a quotient.

  \subsection{Models}

  So, the first stepping stone in talking formally about models are algebraic theories. These are structures with variables and some substitution operation.

  One example of this is the $ \Lambda $-calculus.

  Another example is a polynomial ring. It has variables, you can add and multiply polynomials together. And if you have two polynomials, you can substitute one in the place of a variable in the other to get a new polynomial.

  The formal definition of an algebraic theory is

  \subsection{Semantics}

  \section{The main theorems}

  \subsection{Scott's representation theorem}

  \subsection{The fundamental theorem of the $\lambda$-calculus}

  \subsection{The category of retracts}

  \section{My contribution}

  \subsection{Annotating the paper}

  \subsection{Mechanization}

  \section{Conclusion}

\end{document}