### 2024-01-16 report/chapters/univalent-foundations.tex

\chapter{Univalent Foundations}

\section{Equality and homotopy}
\section{hProps and hSets}
\section{Transports and transport hell}
\section{The univalence axiom}
\section{Univalent categories}