### 2024-02-29 report/report.tex

\documentclass[a4paper]{amsbook}

\usepackage{tikz-cd}
\usepackage{enumerate}
\usepackage{imakeidx}
\usepackage{hyperref}
\usepackage{stmaryrd}

\title{Semantics for the $ \lambda $-calculus}

\makeindex

\newcommand{\iindex}[1]{\textit{#1}\index{#1}}

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}

\theoremstyle{definition}
\newtheorem{definition}{Definition}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}

\newcommand{\id}[1]{\ensuremath{\text{id}_{#1}}}
\newcommand{\op}[1]{\ensuremath{#1^{\text{op}}}}
\newcommand{\Lan}[2]{\ensuremath{\text{Lan}_{#1}#2}}
\newcommand\SET{\mathbf{Set}}
\newcommand\TYPE{\mathbf{Type}}
% \newcommand\F{\mathbf{F}}
\newcommand\FinSET{\mathbf{FinSET}}
\newcommand\AlgTh{\mathbf{AlgTh}}
\newcommand\LamTh{\mathbf{LamTh}}
\newcommand\Alg[1]{\mathbf{Alg}_{#1}}
\newcommand\Pshf[1]{\mathbf{Pshf}_{#1}}
\newcommand\RAct[1]{\mathbf{RAct}_{#1}}
\newcommand\TODO{\textbf{(TODO)} }

\tikzset{%
    symbol/.style={%
        draw=none,
        every to/.append style={%
            edge node={node [sloped, allow upside down, auto=false]{$#1$}}}
    }
}

\labelformat{section}{\thechapter.\thesection}

\begin{document}

  \maketitle

  \setcounter{tocdepth}{4}
  \tableofcontents

  \include{chapters/category-theory}
  \include{chapters/univalent-foundations}
  \include{chapters/algebraic-structures}
  \include{chapters/previous-work}
  \include{chapters/paper}
  \include{chapters/formalization}

  \bibliographystyle{alpha}
  \bibliography{citations}

  \appendix
  \include{chapters/alternative-definitions}

  % \printindex

\end{document}