### 2024-05-29 report/src/paper.tex
Define the theory of extensions

\chapter{Hyland's paper}

\section{Scott's Representation Theorem}
\begin{theorem}\label{thm:representation-theorem}
  Any $ \lambda $-theory $ L $ is isomorphic to the endomorphism $ \lambda $-theory $ E(L) $ of $ L: \Pshf L $, which is $ L $, viewed as a presheaf in its own presheaf category.
\end{theorem}
\begin{proof}
  First of all, remember that $ L $ is indeed exponentiable and that $ L^L = A(L, 1) $.
  Now, since $ L $ is a $ \lambda $-theory, we have sequences of functions back and forth $ \lambda_n: A(L, 1)_n \to L_n $ and $ \rho_n: L_n \to A(L, 1)_n $. These commute with the $ L $-actions, so they constitute presheaf morphisms and $ E(L) $ is indeed a $ \lambda $-theory.

  Lemma \ref{lem:presheaf-Yoneda} gives a sequence of bijections $ \varphi_n : \Pshf L(L^n, L) \cong L_n $ for all $ n $, sending $ F: \Pshf L(L^n, L) $ to $ F(x_1, \dots, x_n) $, and conversely sending $ s: L_n $ to $ ((t_1, \dots, t_n) \mapsto s \bullet (t_1, \dots, t_n)) $. It considers $ \lambda $-terms in $ n $ variables as $ n $-ary functions on the $ \lambda $-calculus. Therefore, it should come as no surprise that $ \varphi $ preserves the $ x_i $, $ \bullet $, $ \rho $ and $ \lambda $, which makes it into an isomorphism of $ \lambda $-theories and this concludes the proof.
\end{proof}

\section{Relations between the categories}

As mentioned before, we have a fully faithful embedding of $ \R $ into $ \Pshf L $.

Also, we have a functor from the Lawvere Theory $ \mathbf{L} $ (see Lemma \ref{lem:lawvere-clone}) associated to $ L $ into $ \R $, sending $ n : \mathbf L $ to $ U^n : \R $. By Scott's representation theorem, this functor can map morphisms as follows:
\[ \mathbf L(m, n) = L_m^n \cong \R(U^m, U)^n \cong \R(U^m, U^n), \]
which immediately shows that this functor is a fully faithful embedding.

Note that if we consider $ L_1 $ as a monoid, with operation $ \bullet $ and unit $ x_1 $, and if we consider this monoid as a category, we can embed this (fully faithfully) into $ \mathbf L $. This gives the following sequence of embeddings:
\begin{center}
  \begin{tikzcd}
    L_1 \arrow[r, hookrightarrow] & \mathbf L \arrow[r, hookrightarrow] & \R \arrow[r, hookrightarrow] & \Pshf L
  \end{tikzcd}
\end{center}

Note that the composition of the first two morphisms sends the object of the category $ L_1 $ to $ (\lambda x_1, x_1) : \R $, which is exactly the usual embedding of $ L_1 $ into its Karoubi envelope.

Write $ i: \op{L_1} \hookrightarrow \op{\mathbf L} $ and $ j: \op{\mathbf L} \hookrightarrow \op \R $ for the embeddings on the opposite categories. By Corollary \ref{cor:surjective-precomposition}, the first two functors in this sequence essentially yield surjective functors on the presheaf categories:
\begin{center}
  \begin{tikzcd}
    C_{L_1} \arrow[d, "Y", hook] &
    \mathbf L \arrow[l, "\op i"'] \arrow[d, "Y", hook] &
    \mathbf R \arrow[rd, hook] \arrow[d, "Y", hook] \arrow[l, "\op j"']\\
    P C_{L_1} &
    P \mathbf L \arrow[l, "i_*"', two heads] &
    P \R \arrow[l, "j_*"', two heads] \arrow[ll, "\sim"', bend left] &
    \Pshf L \arrow[ll, "\sim", bend left]
  \end{tikzcd}
\end{center}
The two equivalences in this diagram are due to Lemma \ref{lem:lawvere-presheaf} and Corollary \ref{cor:karoubi-presheaf}.

\begin{lemma}\label{lem:equivalent-presheaf-cats}
  The precomposition functors $ i_* $ and $ j_* $ are adjoint equivalences.
\end{lemma}
\begin{proof}
  We will show that $ j_* $ is an adjoint equivalence. From the `two out of three' property, it follows then that $ i_* $ is also an adjoint equivalence.

  Lemma \ref{lem:lan-precomp-iso} shows that $ \eta: \id{P \mathbf L} \Rightarrow \Lan j {} \bullet j_* $ is a natural isomorphism. To complete the proof that $ j_* $ is an adjoint equivalence, we just have to show that $ \epsilon: j_* \bullet \Lan j {} \Rightarrow \id{P \mathbf L} $ is a natural isomorphism as well.

  To this end, take $ F : P \R $. By Lemma \ref{lem:lan-precomp-iso}, we have the isomorphism
  \[ \eta^{-1}_{j_* F}: j_* (\Lan j (j_* F)) \cong j_* F. \]
  Since functors preserve isomorphisms, we have $ i_* (j_* (\Lan j (j_* F))) \cong i_* (j_* F) $. However, since $ j_* \bullet i_* $ is an equivalence and in particular fully faithful, this corresponds to an isomorphism
  \[ \Lan j (j_* F) \cong F. \]
  Now we only need to prove that its morphism is equal to $ \epsilon_F $. Equivalently, we need to prove that $ i_* (j_* \epsilon) $ is equal to the morphism $ i_* \eta_{j_* F}^{-1} $, or that $ j_* \epsilon $ equals $ \eta^{-1}_{j_* F} $.

  Take $ n : \mathbf L $. We need to show that $ \epsilon_F (j(n)) = \eta_{j_* F}^{-1}(n) $ as functions from $ (\Lan j {} (j_* F)) (j(n)) $ to $ F(j(n)) $. Note that the diagram for $ (\Lan j {} (j_* F)) (j(n)) $ consists of $ F(j(m)) $ for all $ f: \R(j(m), j(n)) $. Now, as it turns out, both $ \epsilon_F (j(n)) $ and $ \eta_{j_* F}^{-1}(n) $ are defined as the colimit arrow of $ F(f) : \SET(F(j(m)), F(j(n))) $ for all $ f: \R(j(m), j(n)) $, which concludes the proof.
\end{proof}

\begin{lemma}
  The embeddings of $ \R $ into $ \Pshf L $ and $ P \R $ commute with the equivalence between these categories.
\end{lemma}
\begin{proof}
  Note:
  \[ (Y \bullet j_*)(A)(n) = (j \bullet (Y(A)))(n) = Y(A)(j(n)) = \R(U^n, A) \]
  \[ (\iota \bullet \sim)(A)(n) = \{ a : L_n \mid \iota_{0, n}(A) a = a \} \]

  By Scott's representation theorem, we have $ \varphi_n: \R(U^n, U) \xrightarrow \sim L_n, $ given by
  \[ \varphi_n(f) = \iota_{0, n}(f) (x_i)_i. \]
  Restricting this equivalence to the morphisms to $ A $ yields
  \[ \R(U^n, A) \cong \{ a : L_n \mid A a = a \}. \]
  Of course, this is natural in $ A $, since for $ g: \R(A, B) $, postcomposing elements of $ \R(U^n, A) $ by $ g $ is equivalent to applying $ g $ to the elements of $ \{ a : L_n \mid A a = a \} $.

  Lastly, for $ f : L_m^n $, the presheaf actions
  \[ (Y \bullet j_*)(A)(f) : \R(U^n, A) \to \R(U^m, A) \]
  and
  \[ (\iota \bullet \sim)(A)(f) : \{ a : L_n \mid \iota_{0, n}(A) a = a \} \to \{ a : L_m \mid \iota_{0, n}(A) a = a \} \]
  are in fact given by the substitution operations $ \bullet $ in $ E(U) $ and $ L $. Since Scott's representation theorem shows that $ E(U) $ is isomorphic to $ L $ as a $ \lambda $-theory, these substitutions are compatible with the $ \varphi_n $, which shows naturality in $ n $.
\end{proof}

\section{Locally cartesian closedness of the category of retracts}

Recall that Paul Taylor constructed a category of `dependent retracts' $ \R^A $, and then defined a class of display maps for $ \R $ in such a way that it gives a weak equivalence $ \sum_A : \R^A \xrightarrow{\sim} (\R \downarrow_D A) $. I.e., a map $ f: \R(X, A) $ is defined to be a display map if it is isomorphic to something in the image of the functor $ \sum_A $. We saw that these display maps can also be characterized as the retracts of the projection $ p_1: A \times U \to A $ in $ (\R \downarrow A) $.

Now, noticing that we can embed $ \R $ as a full subcategory in $ \Pshf L $, Hyland attempts to do a similar thing for $ \Pshf L $. Noting that the universal object $ U : \R $ is embedded as the theory presheaf $ L : \Pshf L $, he defines a full subcategory $ \R(A) \subseteq (\Pshf L \downarrow A) $, consisting precisely of the retracts of $ \Delta_A L $ in $ (\Pshf L \downarrow A) $.

\begin{remark}
  Again, being a retract of $ \Delta_A L $ is not a mere proposition. For example, consider $ \id L : D(L, L) $:
  \begin{center}
    \begin{tikzcd}
      L \arrow[rr, shift left, "s"] \arrow[rd, "\id L"'] & & L \times L \arrow[ll, shift left, "r"] \arrow[ld, "p_1"]\\
      & L
    \end{tikzcd}
  \end{center}
  There are multiple possible retraction-section pairs $ (r, s) $ that make the diagram commute. For example, we can take $ s $ to be the diagonal embedding $ \Delta : L \to L \times L $ (sending $ l $ to $ (l, l) $). Then we can take $ r = p_1 $ or $ r = p_2 $. We can also take $ r = \pi_1 $ and $ s = \langle \id L, g \rangle $ the product morphism of $ \id L $ and any (possibly constant) endomorphism $ g: L \to L $.

  Therefore, we need to take the propositional truncation of the existence of $ r $ and $ s $.
\end{remark}

\begin{remark}
  Taylor would write $ (\Pshf L \downarrow_D A) $ instead of $ \R(A) $, with $ D(X, A) $ the retracts $ f: X \to A $ of $ \Delta_A L $. However, note that $ D $ is not a class of display maps. For example, for $ D $ to be a class of display maps, all terminal projections $ ! : X \to T $ need to be in $ D $. Since for any category $ C $ with terminal object $ T $, we have an equivalence $ (C \downarrow T) \cong C $, having all terminal projections in $ D $ is equivalent to every object in $ \Pshf L $ being a retract of $ L $.

  However, if we take $ X $ to be the initial (empty) presheaf, we see that there are no morphisms $ L \to X $, because the $ L_n $ are nonempty and the $ X_n $ are empty.

  There also is a more category-theoretical argument: Suppose that every object of $ \Pshf L $ is a retract of $ L $. Via the equivalence $ \Pshf L \cong L_1 $, we see that every object of $ P C_{L_1} $ is a retract of $ Y(\star) $. However then the embedding of $ \hat \R $ (Definition \ref{def:karoubi'}) into $ P C_{L_1} $ is an adjoint equivalence. Since $ Y_{\R} \bullet {\op{\iota_{L_1}}}_* : \R \to \hat \R $ and $ {\op{\iota_{L_1}}}_* : P \R \to P C_{L_1} $ are weak equivalences, (Lemma \ref{lem:retracts-rezk}, Corollary \ref{cor:karoubi-presheaf}), $ Y_\R : \R \to P \R $ is a weak equivalence as well, but the Yoneda embedding is never essentially surjective. For any category $ C $, the constant empty presheaf $ E: c \mapsto \emptyset $ is not representable, since for all $ c : C $, $ \id c = Y(c)(c) $, so $ Y(c) $ is not isomorphic to $ E $.
\end{remark}

Recall that $ \Pshf L $ is isomorphic to the presheaf category $ P \mathbf L $, so it is an elementary topos, so it is locally cartesian closed (Example 5.2.5 and Theorem 5.8.4, \autocite{borceux}, Volume 3). Therefore, for a morphism $ f: \Pshf L(Y, X) $, we have adjunctions
\begin{center}
  \begin{tikzcd}
    (\Pshf L \downarrow X) \arrow[r, "f^*"' description, ""{name=B}] & (\Pshf L \downarrow Y) \arrow[l, bend right, "{\sum_f = - \cdot f}"'{name=A}] \arrow[l, bend left, "\prod_f"{name=C}]
    \arrow[from=A, to=B, symbol=\dashv]
    \arrow[from=B, to=C, symbol=\dashv]
  \end{tikzcd}
\end{center}

Now, Hyland shows the following:
\begin{theorem}\label{thm:restrict-sum-product}
  For $ (Y, f) : \R(X) $, we can restrict $ \sum_f $ and $ \prod_f $ to functors from $ \R(Y) $ to $ \R(X) $.
\end{theorem}
\begin{proof}
  This proof proceeds by reducing, step by step, to a known case.

  We take $ (Z, g) : \R(Y) $. Since functors preserve retraction-section pairs, $ \prod_f(Z, g) $ is a retract of $ \prod_f \Delta_Y L $. Now, if this is a retract of $ \Delta_X L $, then $ \prod_f(Z, g) $ is a retract of $ \Delta_X L $ as well: We can just compose retractions and sections. Therefore, to show that the image of $ \R(Y) $ under $ \prod_f $ lies within $ \R(X) $, we just need to show that $ \prod_f \Delta_Y L $ lies in $ \R(X) $. In the same way, for $ \sum_f $ we just need to show that $ \sum_f \Delta_Y L $ lies in $ \R(X) $.

  By the definition of $ \R(X) $, $ (Y, f) $ is a retract:
  \begin{center}
    \begin{tikzcd}
      Y \arrow[rd, "f"'] \arrow[r, "s"] & X \times L \arrow[d, "p_1"] \arrow[r, "r"] & Y \arrow[ld, "f"]\\
      & X
    \end{tikzcd}
  \end{center}

  The counits of the adjunctions $ r_* \vdash \prod_r $ and $ s_* \vdash \prod_s $, give maps
  \[
    \prod_f \Delta_Y L \xrightarrow{\prod_f \eta_r}
    \prod_f \prod_r r^* \Delta_Y L \xrightarrow{\prod_f \prod_r \eta_s}
    \prod_f \prod_r \prod_s s^* r^* \Delta_Y L
  \]
  Their composite is $ \prod_f \eta_{s \cdot r} $. However, note that $ r \cdot s = \id Y $, so we have an isomorphism of functors
  \[ (s \cdot r)^* \cong \id{\Pshf L \downarrow Y} \quad \text{and} \quad \prod_{s \cdot r} \cong \id{\Pshf L \downarrow Y} \]
  and transporting $ \eta_{s \cdot r} $ along these isomorphisms, we get the identity natural transformation $ \id{\Pshf L \downarrow Y} \Rightarrow \id{\Pshf L \downarrow Y} $.

  Secondly, note that $ r \cdot f = p_1 $, so we have $ \prod_r \bullet \prod_f \cong \prod_{p_1} $.

  Lastly, by Remark \ref{rem:pullback-of-projection}, $ r^* \Delta_Y L \cong \Delta_{X \times L} L $.

  Therefore, we have the following diagram, showing that $ \prod_f \Delta_Y L $ is a retraction of $ \prod_{p_1} \Delta_{X \times L} L $:
  \begin{center}
    \begin{tikzcd}
      & \prod_{p_1} \Delta_{X \times L} L\\
      \prod_f \Delta_Y L
        \arrow[r, "\prod_f \eta_r"]
        \arrow[rr, bend right, "\prod_f \eta_{s \cdot r}"]
        \arrow[rrr, bend right, "\id {\Pshf L \downarrow X}"] &
      \prod_f \prod_r r^* \Delta_Y L \arrow[r, "\prod_f \prod_r \eta_s"] \arrow[u, "\sim"] &
      \prod_f \prod_r \prod_s s^* r^* \Delta_Y L \arrow[r, "\sim"] &
      \prod_f \Delta_Y L
    \end{tikzcd}
  \end{center}
  In a similar way, $ \sum_f \Delta_Y L $ is a retract of $ \sum_{p_1} \Delta_{X \times L} L $:
  \begin{center}
    \begin{tikzcd}
      & & \sum_{p_1} \Delta_{X \times L} L \\
      \sum_f \Delta_Y L
        \arrow[r, "\sim"]
        \arrow[rrr, bend right, "\id {\Pshf L \downarrow X}"] &
      \sum_f \sum_r \sum_s s^* r^* \Delta_Y L
        \arrow[r, "\sum_f \sum_r \epsilon_s"]
        \arrow[rr, bend right, "\sum_f \epsilon_{s \cdot r}"] &
      \sum_f \sum_r r^* \Delta_Y L
        \arrow[u, "\sim"]
        \arrow[r, "\sum_f \epsilon_r"] &
      \sum_f \Delta_Y L
    \end{tikzcd}
  \end{center}
  So again, by the transitivity of retracts, we only have to show that $ \prod_{p_1} \Delta_{X \times L} L $ and $ \sum_{p_1} \Delta_{X \times L} L $ are retracts of $ \Delta_X L $. By Lemma \ref{lem:constant-dependent-product}, we have $ \prod_{p_1} \Delta_{X \times L} L \cong \Delta_X L^L $. By Lemma \ref{lem:sum-postcomposition}, the functor $ \sum_{p_1} $ is given by postcomposition, so $ \sum_{p_1} \Delta_{X \times L} L \cong \Delta_X (L \times L) $.

  Since $ \Delta_X $ is a functor, and in particular, preserves retracts, we only have to show that $ L \times L $ and $ L^L $ are retracts of $ L $.

  But we already saw in Theorem \ref{thm:representation-theorem} that $ L^L \cong A(L, 1) $ is a retract of $ L $.

  Also, the (pairing and splitting) morphisms $ s: \Pshf L(L \times L, L) $ and $ r: \Pshf L(L, L \times L) $ given by
  \[ s_n(a, b) = \lambda x_{n + 1}, x_{n + 1} a b \quad \text{and} \quad r_n(a) = (a (\lambda x_{n + 1} x_{n + 2}, x_{n + 1}), a (\lambda x_{n + 1} x_{n + 2}, x_{n + 2})) \]
  exhibit $ L \times L $ as a retract of $ L $.
\end{proof}

\begin{remark}
  Now, recall that according to Hyland, Taylor endows the category of retracts $ \R $ with a class of display maps $ D $, such that $ (\R \downarrow_D X) $ consists exactly of the retracts of $ \Delta_X U $.

  Hyland then claims that the fact that $ \R $ is cartesian closed relative to this class of display maps follows from the theorem above. Recall that Taylor already shows relative cartesian closedness directly, which still works in univalent foundations if we assume the axiom of choice.

  Recall that we have a weak equivalence $ \varphi : \R \xrightarrow \sim \hat \R $ (Lemma \ref{lem:retracts-rezk}) and one can imagine that we can extend this to weak equivalences $ \psi_X : (\R \downarrow_D X) \xrightarrow \sim (\hat \R \downarrow_D \varphi(X)) $. Hyland notes that for $ X : \hat \R $, $ \R(\varphi(X)) $ is equivalent to $ (\hat \R \downarrow_D \varphi(X)) $. Implicitly, for $ f: \R(X, Y) $, he wants to lift the functors $ \sum_f $, $ f^* $ and $ \prod_f $ along the equivalences $ \psi $:
  \begin{center}
    \begin{tikzcd}[sep=large]
      (\R \downarrow_D X)
        \arrow[d, "\psi_X"]
        \arrow[r, dashed, bend left, "\sum_f"']
        \arrow[r, dashed, bend right, "\prod_f"]
      & (\R \downarrow_D Y)
        \arrow[d, "\psi_Y"]
        \arrow[l, dashed, "f^*" description]\\
      \R(\varphi(X))
        \arrow[r, bend left, "\sum_{\varphi(f)}"']
        \arrow[r, bend right, "\prod_{\varphi(f)}"]
      & \R(\varphi(Y))
        \arrow[l, "\varphi(f)^*" description]\\
    \end{tikzcd}
  \end{center}
  For this to work, we need to turn the weak equivalences $ \psi_X $ into adjoint equivalences. In particular, we know that for every object $ A : \R(\varphi(X)) $, there merely exists a preimage $ \bar A : (\R \downarrow_D X) $ such that $ \psi_X(\bar A) \cong A $, and we need to construct a functor $ \psi_X^{-1}: \R(\varphi(X)) \to (\R \downarrow_D X) $.

  For Taylor's proof, we did a similar thing when we knew that for every object $ X : (\R \downarrow_D A) $, there existed some $ Y : \R^A $, and we needed to construct a pullback functor $ f^* : (\R \downarrow_D A) \to (\R \downarrow_D B) $. We could use the axiom of choice here because the objects of $ \R $ (and therefore also the objects of the relative slices $ (\R \downarrow_D A) $) form a set.

  However, as mentioned in \autocite{univalent-categories} halfway through the introduction, and as explained more thoroughly in \autocite{hottbook}, Lemma 3.8.5, if $ \R(\varphi(X)) $ is not a set (and if $ L $ is nontrivial, it is never a set), then there is no axiom of choice which can turn the `for all \dots, there exists \dots' into a `there exists a function'.

  Therefore, in univalent foundations, the relative cartesian closedness is not a corollary of Theorem \ref{thm:restrict-sum-product}.
\end{remark}

\section{An elementary proof of the `Fundamental Theorem'}

The final theorem that Hyland works towards in his paper constructs an equivalence between $ \lambda $-calculi and $ \Lambda $-algebras. In this section, we will provide an elementary proof of this theorem.

We will assume that $ \Lambda $ (and therefore, any $ \lambda $-theory) satisfies $ \beta $-equality.

In both cases, one part of the equivalence is very easy:
\begin{definition}
  For all $ n $, we have a functor from lambda theories to $ \Lambda $-algebras. It sends the $ \lambda $-theory $ L $ to the $ L $-algebra $ L_n $ and then turns this into a $ \Lambda $-algebra via the morpism $ \Lambda \to L $. It sends morphisms $ f: L \to L^\prime $ to the algebra morphism $ f_n : L_n \to L^\prime_n $.
\end{definition}
Here we are only interested in the case $ n = 0 $.

We will use Lemma \ref{lem:univalen-category-equivalence}, by showing that $ L \mapsto L_0 $ is a weak equivalence.

First of all, let $ A $ be an algebra for the initial $ \lambda $-theory $ \Lambda $.

The $ \Lambda $-algebra structure gives the terms of $ A $ quite a lot of interesting behaviour. For example, we can define `function application' and composition as
\[ a b = (x_1 x_2) \bullet (a, b) \quad \text{and} \quad a \circ b = (x_1 \circ x_2) \bullet (a, b), \]
and the same for the other constructions at the start of Section \ref{sec:retracts-category}.

\begin{remark}
  Recall that in Example \ref{ex:free-monoid-theory}, we constructed an algebraic theory $ T $ encapsulating the structure of a monoid. This allowed us to define a monoid operation on $ T $-algebras as well. We then were able to transfer associativity of the operation on the $ T_n $ to associativity of the operation on the algebras. In exactly the same way, the function composition on $ A $ is associative because composition on $ \Lambda_n $ is associative. Similarly, we can show that $ \pi_1 (a, b) = a $, that $ \pi_2 \circ \langle a, b \rangle = (\lambda x_2, x_1 x_2) \bullet b $ etc.
\end{remark}

In fact, we can repeat almost the entirety of Section \ref{sec:retracts-category} for $ A $ instead of for $ L_0 $:
\begin{definition}
  We define the `category of retracts' of $ A $ as the category $ \R_A $ given by
  \[ (\R_A)_0 = \{ X : A \mid X \circ X = X \} \quad \text{and} \quad \R_A(X, Y) = \{ f : A \mid Y \circ f \circ X = f \}. \]
\end{definition}
Just like in Section \ref{sec:retracts-category}, $ \R_A $ has `universal object' and terminal object
\[ U = (\lambda x_1, x_1) \bullet () \quad \text{and} \quad I = (\lambda x_1, c_1), \]
products
\[ A \times B = \langle A \circ \pi_1, B \circ \pi_2 \rangle \]
and exponential objects
\[ B^A = (\lambda x_3, x_1 \circ x_3 \circ x_2) \bullet (B, A) \]
with the isomorphism $ \psi: \R_A(C \times A, B) \xrightarrow \sim \R_A(C, B^A) $
given by
\[ \psi(f) = (\lambda x_2 x_3, x_1 (x_2, x_3)) \bullet f \quad \text{and} \quad \psi^{-1}(f) = (\lambda x_2, x_1 (\pi_1 x_2) (\pi_2 x_2)) \bullet f. \]

Therefore, we have a $ \lambda $-theory $ E_{\R_A}(U) $.

\begin{remark}
  Note that if $ A = L_0 $, then the construction of $ \R_A $ and $ \R $ coincide, so we have an almost trivial equivalence $ \R_A \cong \R $, and an isomorphism $ E_{\R}(U) \cong E_{\R_A}(U) $.
\end{remark}

The following lemma shows that our functor is essentially surjective:

\begin{lemma}\label{lem:elementary-essentially-surjective}
  We have a $ \Lambda $-algebra isomorphism $ \epsilon_A : E_{\R_A}(U)_0 \xrightarrow \sim A $.
\end{lemma}
\begin{proof}
  Take $ c_n = (\lambda x_{n + 1}, x_{n + 1}): \Lambda_n $. Note that
  \[ E(U)_0 = \{ f : A \mid (\lambda x_2, x_1 c_2) \bullet f = f \}. \]
  Therefore, we have a bijection given by
  \[ \epsilon(a) = (x_1 c_1) \bullet a \quad \text{and} \quad \epsilon^{-1}(a) = (\lambda x_2, x_1) \bullet a. \]
  Now, to show that this is an isomorphism of $ \Lambda $-algebras, recall that $ \bullet: \R_A(U^n, U) \times \R_A(I, U)^n \to \R_A(U) $ is given by precomposition with the product morphism and that the $ \Lambda $-algebra structure on $ E(U)_0 $ is given by the embedding of $ \Lambda $ into $ E(U)_0 $, given by $ f \mapsto {\iota_\Lambda}_n(f) \bullet_A () $ for some collection $ ({\iota_\Lambda}_n: \Lambda_n \to \Lambda_0)_n $ (from now on, we will drop the $ n $ and just write $ \iota_\Lambda $). We have for all $ f: \Lambda_n $, and all $ a : E(U)_0^n $,
  \begin{align*}
    \epsilon(f \bullet_{E(U)} a)
    &= ((\iota_\Lambda(f) \bullet_A ()) \circ \langle a_i \rangle_i)(c_n)\\
    &= (\iota_{0, n}(\iota_\Lambda(f)) (x_i c_n)_i) \bullet_A a\\
    &= (\iota_{0, n}(\iota_\Lambda(f)) (x_i)_i) \bullet_A (\epsilon(a_i))_i,
  \end{align*}
  and we want this to equal $ f \bullet_A (\epsilon(a_i))_i $. Leaving out the $ \iota_{0, n} $, all we need to do is show that for all $ f: \Lambda_n $, we have $ \iota_\Lambda(f) (x_i)_i = f $. We will do this by structural induction on $ f $.
  \begin{itemize}
    \item If $ f = x_{n, i} $, we have $ \iota_\Lambda(f) = \pi_{n, i} $, and
      \[ \iota_\Lambda(f) (x_i)_i = \pi_{n, i} (x_{n, i})_i = x_{n, i}. \]
    \item If $ f = \lambda x_{n + 1}, g $ for some $ g: \Lambda_{n + 1} $, we have
      \[ \iota_\Lambda(f) = (\lambda x_1 x_2, \iota_{0, 2}(\iota_\Lambda(g)) (x_1, x_2)), \]
      and if the induction hypothesis holds for $ g $, then
      \begin{align*}
        \iota_\Lambda(f) (x_i)_i
        &= (\lambda x_{n + 1} x_{n + 2}, \iota_\Lambda(g) (x_{n + 1}, x_{n + 2})) (x_i)_i\\
        &= \lambda x_{n + 1}, \iota_\Lambda(g) ((x_i)_i, x_{n + 1})\\
        &= \lambda x_{n + 1}, g.
      \end{align*}
    \item Recall that application in a $ \lambda $-theory is given by $ g_1 g_2 = \rho(g_1) \bullet (x_1, \dots, x_n, g_2) $. Now, if $ f = g_1 g_2 $ for $ g_1, g_2 : \Lambda_{n + 1} $, we have
      \[
        \iota_\Lambda(f)
        = (\lambda x_1, \iota_\Lambda(g_1) (\pi_1 x_1) (\pi_2 x_1)) \circ \langle \id{U^n}, \iota_\Lambda(g_2) \rangle
        = \lambda x_1, \iota_\Lambda(g_1) x_1 (\iota_\Lambda(g_2) x_1),
      \]
      and if the induction hypothesis holds for $ g_1 $ and $ g_2 $, then
      \begin{align*}
        \iota_\Lambda(f) (x_i)_i
        &= (\lambda x_{n + 1}, \iota_\Lambda(g_1) x_{n + 1} (\iota_\Lambda(g_2) x_{n + 1})) (x_i)_i\\
        &= \iota_\Lambda(g_1) (x_i)_i (\iota_\Lambda(g_2) (x_i)_i)\\
        &= g_1 g_2.
      \end{align*}
  \end{itemize}
\end{proof}

Now, take a morphism $ F: \Alg{\Lambda}(A, B) $. Recall that $ \circ $, the categories $ \mathbf R_A $ and $ \mathbf R_B $, and their products and exponential objects are given using $ f \bullet (a_1, \dots, a_n) $ for some $ f: \Lambda_n $ and $ a_1, \dots, a_n : A $. Since we have $ F(f \bullet (a_i)_i) = f \bullet (F(a_i)) $, $ F $ gives a functor $ F: \R_A \to \R_B $, such that
\[ F(I) = I, F(U) = U, F(A \times B) = F(A) \times F(B) \quad \text{and} \quad F(A^B) = F(A)^{F(B)}, \]
and the same for the product projections and the the natural isomorphisms $ \R_A(C, B^A) \cong \R_A(C \times A, B) $. In particular, we have for all $ f: \R_A(U^n, U) $,
\[ F(f) : \R_B(F(U^n), F(U)) = \R_B(U^n, U). \]
Therefore, $ F $ gives a $ \Lambda $-theory morphism between the endomorphism theories of $ U: \R_A $ and $ U: \R_B $. This allows us to show:
\begin{lemma}\label{lem:elementary-full}
  The functor is full.
\end{lemma}
\begin{proof}
  For any $ \lambda $-theory $ L $, Theorem \ref{thm:Scott} gives an isomorphism of $ \lambda $-theories $ \eta_L: L \xrightarrow \sim E_{\R_{L_0}}(U) $, with
  \[ {\eta_L}_0(f) = \lambda x_1, \iota_{0, 1}(f) \quad \text{and} \quad {\eta_L^{-1}}_0(g) = g(\lambda x_1, x_1). \]

  Now, for $ L, L^\prime: \LamTh $ and $ F: \Alg{\Lambda}(L_0, L^\prime_0) $, we have
  \[ \eta_L \cdot F \cdot \eta_{L^\prime}^{-1} : \LamTh(L, L^\prime), \]
  and we have for all $ s : L_0 $,
  \[ ({\eta_L}_0 \cdot F \cdot {\eta_{L^\prime}^{-1}}_0)(s) = (\lambda x_1, \iota_{0, 1}(F(s))) (\lambda x_1, x_1) = F(s), \]
  which shows that the functor is full.
\end{proof}

Now, we only have to show:
\begin{lemma}\label{lem:elementary-faithful}
  The functor is faithful.
\end{lemma}
\begin{proof}
  Take morphisms $ F, G: \LamTh(L, L^\prime) $. Suppose that $ F_0 = G_0 $. Then, for all $ s: L_n $, we have
  \[ F_n(s) = \rho^n(F_0(\lambda^n(s))) = \rho^n(G_0(\lambda^n(s))) = G_n(s), \]
  so $ F = G $.
\end{proof}

Summarizing,
\begin{theorem}
  The functor that sends $ L : \LamTh $ to $ L_0 : \Alg{\Lambda} $, is an adjoint equivalence.
\end{theorem}
\begin{proof}
  By Lemma \ref{lem:elementary-full} and Lemma \ref{lem:elementary-faithful}, the functor is fully faithful. By Lemma \ref{lem:elementary-essentially-surjective}, it is essentially surjective, so it is a weak equivalence. Since $ \LamTh $ and $ \Alg{\Lambda} $ are univalent categories, Lemma \ref{lem:univalen-category-equivalence} shows that $ F $ is an adjoint equivalence.
\end{proof}

\section{Hyland's proof}

Hyland gives a more category theoretical proof. That means that there is more high-level intuition why things work the way they work, but on the flip side, there is a lot more details to check.

\subsection{Terms of a \texorpdfstring{$ \Lambda $}{Lambda}-algebra}

\begin{definition}
  Take $ \mathbf 1_n = (\lambda x_1 \dots x_n, x_1 \dots x_n) \bullet () : A $.
\end{definition}

\begin{remark}
  Note that we have $ \mathbf 1_1 = U : \R_A $.
\end{remark}

\begin{definition}
  In this section, we consider sets of elements of $ A $ that behave like functions in $ n $ variables:
  \[ A_n = \{ a : A \mid (\lambda x_2 x_3 \dots x_{n + 1}, x_1 x_2 x_3 \dots x_{n + 1}) \bullet a = a \}. \]
\end{definition}

\begin{remark}
  Some straightforward rewriting, shows that
  \[ A_n = \{ a : A \mid \mathbf 1_n \circ a = a \}. \]
\end{remark}

\begin{remark}
  Also note that $ \mathbf 1_n \circ a \circ \mathbf 1_n = \mathbf 1_n \circ a $, so for $ a : A_n $, $ a \circ \mathbf 1_n = a $.
\end{remark}

The following shows that `functions in $ n $ variables' are indeed all in $ A_n $:
\begin{lemma}
  For $ t: \Lambda_{m + n} $ and $ a_1, \dots, a_m: A $, we have $ (\lambda^n t) \bullet (a_1, \dots, a_m) : A $ and we have
  \[ \mathbf 1_n \circ ((\lambda^n t) \bullet (a_1, \dots, a_m)) = (\lambda^n t) \bullet (a_1, \dots, a_m), \]
  so $ (\lambda^n t) \bullet (a_1, \dots, a_m) : A_n $.
\end{lemma}
\begin{proof}
  This follows by straightforward rewriting.
\end{proof}

\begin{corollary}
  By the previous remark,
  \[ ((\lambda^n t) \bullet (a_1, \dots, a_m)) \circ \mathbf 1_n = (\lambda^n t) \bullet (a_1, \dots, a_m). \]
\end{corollary}
\begin{corollary}
  In particular, $ \mathbf 1_m \circ \mathbf 1_n = \mathbf 1_{\max(m, n)} $. From this, it follows that $ A_m \subseteq A_n $ for $ m \leq n $. It also follows that $ a \mapsto \mathbf 1_n \circ a $ gives a function from $ A $ to $ A_n $ (and also from $ A_m \subseteq A $ to $ A_n $).
\end{corollary}

\begin{definition}[The monoid of a $ \Lambda $-algebra]
  We make $ A_1 $, the `functional elements' of $ A $, into a monoid under composition $ \circ $ with unit $ \mathbf 1_1 $. The fact that this is a monoid follows from the remarks above.
\end{definition}

Recall that we have an equivalence $ [\op{C_{A_1}}, \SET] \cong \RAct{A_1} $.

\begin{remark}\label{rem:ract-presheaf-equivalence}
  Note that, like in the last chapter, $ \R_A $ pops up as the Karoubi envelope of the monoid category $ C_{A_1} $, and fits into the following diagram:
  \begin{center}
    \begin{tikzcd}
      & C_{A_1} \arrow[d, hookrightarrow, "Y"] \arrow[r, hookrightarrow, "\iota_{C_{A_1}}"] & \R_A \arrow[d, hookrightarrow, "Y"]\\
      \RAct{A_1} & P C_{A_1} \arrow[l, "\sim"] & P \R_A \arrow[l, "\sim"]
    \end{tikzcd}
  \end{center}
  Explicitly, this gives the embedding $ \R_A(\mathbf 1_1, -): \R_A \hookrightarrow \RAct{A_1} $ given by
  \[ X \mapsto \R_A(\mathbf 1_1, X) = \{ x : A \mid X \circ x = x \}. \]
  Also, note that if $ A = L_0 $ for some $ \lambda $-theory $ L $, then the monoid $ A_1 $ is equivalent to the monoid $ L_1 $, and $ \R_A $ is equivalent to our familiar category of retracts $ \R $. Using \ref{lem:equivalent-presheaf-cats}, we have the following 2-commutative diagram:
  \begin{center}
    \begin{tikzcd}
      &
      C_{{L_0}_1} \arrow[dd, "Y" description, hook] \arrow[rrr, "\iota_{C_{{L_0}_1}}" description, hook] \arrow[rd, "\sim" description, no head] & & & \R_{L_0} \arrow[dd, "Y" description, hook] \arrow[rd, "\sim" description, no head] & & \\
      &
      & C_{L_1} \arrow[dd, "Y" description, hook] \arrow[r, hook] & \mathbf L \arrow[rr, hook] \arrow[dd, "Y" description, hook] & & \mathbf R \arrow[dd, "Y" description, hook] \arrow[rdd, hook] & \\
      \RAct{{L_0}_1} &
      P C_{{L_0}_1} \arrow[l, "\sim"'] \arrow[rd, "\sim" description, no head] & & & P \R_{L_0} \arrow[lll, "\sim" description] \arrow[rd, "\sim" description, no head] & & \\
      &
      & P C_{L_1} & P \mathbf L \arrow[l, "\sim" description] & & P \R \arrow[ll, "\sim" description] & \Pshf L \arrow[lll, "\sim" description, bend left]
    \end{tikzcd}
  \end{center}
\end{remark}

\subsection{Constructing a theory from an algebra}

\begin{definition}
  Composition $ \circ $ gives a right $ A_1 $-action on the $ A_n $, so we have $ A_n : \RAct{A_1} $. In particular, $ A_1 $ acts on itself, and we will call this set with right $ A_1 $-action $ U_A $.
\end{definition}

\begin{lemma}
  $ A_2 $ is isomorphic to the exponential object $ U_A^{U_A} $ in $ \RAct{A_1} $.
\end{lemma}
\begin{proof}
  Recall that $ U_A^{U_A} $ is the set of $ A_1 $-equivariant morphisms $ U_A \times U_A \to U_A $. We have an isomorphism $ \psi: A_2 \xrightarrow \sim U_A^{U_A} $, given by
  \[ \psi(f) = (b, b^\prime) \mapsto (\lambda x_4, x_1 (x_2 x_4) (x_3 x_4)) \bullet (f, b, b^\prime), \]
  treating $ f : A_2 $ as a function in two variables, and simultaneously composing it with the functions $ b, b^\prime : A_1 $. It has an inverse
  \[ \psi^{-1}(g) = (\lambda x_2 x_3, x_1 (x_2, x_3)) \bullet (g(\pi_1, \pi_2)). \]
  Note that the first pair $ (x_2, x_3) $ is a $ \lambda $-term, whereas the second pair $ (\pi_1, \pi_2) $ is a pair in $ U_A \times U_A $.

  For $ f: U_A^{U_A} $ and $ (a_1, a_2): U_A \times U_A $, we have
  \[
    \psi(\psi^{-1}(f))(a_1, a_2)
    = f(\pi_1, \pi_2) \circ \langle a_1, a_2 \rangle
    = f(\pi_1 \circ \langle a_1, a_2 \rangle, \pi_2 \circ \langle a_1, a_2 \rangle)
    = f(a_1, a_2).
  \]
  Here we use the $ A_1 $-equivariance of $ f $. In the last step of this proof, we use, among other things, the fact that the $ a_i : A_1 $ and therefore $ \lambda x_1, a_i x_1 = a_i $.

  Some straightforward rewriting shows that for $ a: A_2 $, we have $ \psi^{-1}(\psi(a)) = a $. In the last step of this proof, we use the fact that $ a : A_2 $ and therefore $ \lambda x_1 x_2, a x_1 x_2 = a $.

  Therefore, $ \psi $ is a bijection and, as it turns out, an isomorphism.
\end{proof}

\begin{remark}
  Recall that the embedding $ \R_A \hookrightarrow \RAct{A_1} $ is the composition of a Yoneda embedding and two adjoint equivalences. These all preserve exponential objects (\TODO: Citation needed). Now, note that $ A_1 $ is the image of $ U : \R_A $, so the exponential $ U_A^{U_A} $ is $ A_2 $, the image of $ U^U = \lambda x_1 x_2, x_1 x_2 = \mathbf 1_2 $.
\end{remark}

\begin{definition}[Construction of the $ \lambda $-theory]
  Recall that $ A_2 \subseteq U_A $, and that we have a retraction $ a \mapsto \mathbf 1_2 \circ a : U_A \to A_2 $.

  Therefore, $ U_A $ is a reflexive object in $ \RAct{A_1} $, and we get a $ \lambda $-theory $ E(U_A) $.
\end{definition}

\begin{remark}
  The cartesian closed embedding of $ \R_A $ into $ \RAct{A_1} $ sends $ U $ to $ U_A $. It sends the retraction $ U^U: \R_A(U, U^U) $ exactly to the retraction $ a \mapsto \mathbf 1_2 \circ a: \RAct{A_1}(U_A, A_2) $. Therefore, we have a $ \lambda $-theory isomorphism $ E_{\R_A}(U) \cong E_{\RAct{A_1}}(U_A) $.
\end{remark}

\subsection{Constructing a theory morphism from an algebra morphism}

Take a morphism $ F: \Alg{\Lambda}(A, B) $. Note that $ F $ preserves $ \circ $ and the $ \mathbf 1_n $. Therefore, it induces a monoid morphism $ A_1 \to B_1 $, which gives a functor $ C_{A_1} \to C_{B_1} $. Precomposing with this, we get a functor $ \RAct{B_1} \to \RAct{A_1} $.

To get a morphism $ E_{\RAct{A_1}}(U_A) \to E_{\RAct{B_1}}(U_B) $ however, we need a functor $ \RAct{A_1} \to \RAct{B_1} $, which we obtain by Left Kan extension: see Lemma \ref{lem:scalar-extension} (the ``tensor product'').

Hyland gives a very high-level argument why $ F_* $ induces a morphism $ \bar F : \LamTh(E_{\RAct{A_1}}(U_A), E_{\RAct{B_1}}(U_B)) $, which we will discuss a bit more here.

\begin{lemma}
  The extension of scalars $ F_* $ sends $ U_A $ to $ U_A $ and preserves finite products.
\end{lemma}
\begin{proof}
  By Lemma \ref{lem:scalar-extension-monoid-monoid-action}, we have $ F_*(U_A) \cong U_A $.

  We will show that $ F_* $ preserves binary products and the terminal object, because from this it follows that $ F_* $ preserves all finite products.

  We use Lemma \ref{lem:scalar-extension-terminal} to show that $ F_* $ preserves the terminal object. We take (very similar to the terminal object in $ \R $):
  \[ a_0 = (\lambda x_1 x_2, x_2) \bullet () : A_1 \quad \text{and} \quad b_0 = (\lambda x_1 x_2, x_2) \bullet () : B_1 \]
  and $ a_0 $ is weakly terminal because for all $ a : B_1 $, we have $ F(a_0) \circ a = b_0 $.

  We use Lemma \ref{lem:scalar-extension-product} to show that $ F_* $ also preserves the product. Therefore, given $ a_1, a_2 : B_1 $. Take $ b = \langle a_1, a_2 \rangle $, together with the familiar projections $ \pi_i : A_1 $. We have $ a_i = F(\pi_i) \circ a $.

  Now, for some $ b^\prime: B_1 $ and $ \pi_1^\prime, \pi_2^\prime: A_1 $ such that $ a_i = F(\pi_i^\prime) \circ b^\prime $, take $ m = \langle \pi_1^\prime, \pi_2^\prime \rangle $. Then $ \pi_i \circ m = \pi_i^\prime $ and $ F(m) \circ a^\prime = a $, so $ (a, \pi_1, \pi_2) $ is weakly terminal and $ F_* $ preserves binary products.

  Since any finite product is (isomorphic to) a construction with a repeated binary product and the terminal object, the fact that $ F_* $ preserves binary products and the terminal object shows that $ F_* $ preserves all finite products.
\end{proof}

Now, we can start defining our lift of $ F $:
\begin{definition}
  We can send an element $ g: E(U_A)_n = \RAct{A}(U_A^n, U_A) $ to
  \[ F_*(g): \RAct{A^\prime}(f(U_A^n), f(U_A)) \cong \RAct{A^\prime}((U_{A^\prime})^n, U_{A^\prime}) = E(U_{A^\prime})_n \]
  so we have a morphism $ \bar F: \LamTh(E(U_A), E(U_{A^\prime})) $.
\end{definition}

\begin{remark}
  The fact that $ \bar F $ preserves the variables and substitution is not very hard, since these are just defined in terms of finite products of $ U_A $ and $ \bar F $ preserves finite products and $ U_A $.

  However, showing that it is a $ \lambda $-theory morphism is a different matter. Hyland claims

  \enquote{$ F $ preserves $ \mathbf 1_n $, which determines the function space as a retract of the universal. So $ \bar F $ preserves the retract and the result follows.}

  Although this covers the core of the argument, it is very complicated to actually verify that this works, because we need to pass through a lot of isomorphisms, and check that they work nicely together:
  \begin{align*}
    \alpha_A: \RAct{A_1}(X \times Y, Z) &\xrightarrow{\sim} \RAct{A_1}(X, Z^Y);\\
    \beta: F_*(U_A) &\xrightarrow{\sim} U_{A^\prime};\\
    \bar \gamma: F_*(A \times B) &\xrightarrow{\sim} F_*(A) \times F_*(B);\\
    \gamma_n: F_*(X^n) &\xrightarrow{\sim} F_*(X)^n;\\
    % \delta: F_*(X^Y) &\xrightarrow{\sim} F_*(X)^{F_*(Y)};\\
    \delta_A: U_A^{U_A} &\xrightarrow{\sim} A_2.
  \end{align*}
  Since we already motivated this in a different way, we will leave it as an exercise for the enthousiastic reader.
  % with $ \gamma_{n + 1} = \bar{\gamma} \cdot (\gamma_n \times \id X) $.
  % Then, for $ s: E(U_A)_{n + 1} $, we have $ \lambda_A(s) = \alpha_A(s) \cdot \delta_A \cdot \pi_A : \RAct{A_1}(U_A^n, U_A) $ for the projection $ \pi_A: A_2 \to U_A $. However, note that $ f^*(\lambda_A(s)): \RAct{A^\prime_1}(f^*(U_{A^\prime}^n), f^*(U_{A^\prime})) $. To compare this to $ \lambda_{A^\prime}(\dots) $, we actually need to take
  % \[ (\beta^{-1})^n \cdot \gamma_n^{-1} \cdot f^*(\lambda_A(s)) \cdot \beta : \RAct{A^\prime_1}(U_A^n, U_A). \]
  % On the other hand, we also have $ f^*(s): \RAct{A^\prime_1}(f^*(U_A^{n + 1}), f^*(U_A)) $ and we cannot apply $ \lambda_{A^\prime} $ directly. Instead, the term that we end up with is
  % \[ \lambda_{A^\prime}((\beta^{-1})^{n + 1} \cdot \gamma_{n + 1}^{-1} \cdot f^*(s) \cdot \beta): \RAct{A^\prime_1}(U_A^n, U_A). \]
  % So the equality that we need to prove is (using functoriality of $ f^* $ and naturality of $ \alpha_{A^\prime} $):
  % \begin{align*}
  %   &(\beta^{-1})^n \cdot \gamma_n^{-1} \cdot f^*(\alpha_A(s)) \cdot f^*(\delta_A \cdot \pi_A) \cdot \beta\\
  %   &= (\beta^{-1})^n \cdot \gamma_n^{-1} \cdot \alpha_{A^\prime}((\id X^n \times \beta^{-1}) \cdot \bar{\gamma}^{-1} \cdot f^*(s)) \cdot \beta^{U_A} \cdot \delta_{A^\prime} \cdot \pi_{A^\prime}.
  % \end{align*}
  % Because of the complicated definition of each of these terms, I was unable to verify the correctness of this statement (and the similar statement about $ \rho: E(U_A)_n \to E(U_A)_{n + 1} $) within a day, even though I am willing to believe Hyland's claim.

  % Given a morphism $ t: X \times Y \to Z $, $ \alpha_{A_1}(t): X \to (U_A \times Y \to Z) $ is given by
  % \[ \alpha(t)(x)(a, y) = t(x a, y) \]

  % Also, $ \beta: f^*(U_A) \xrightarrow{\sim} U_{A^\prime} $ is given, for $ a: A_1 $ and $ a^\prime : A^\prime_1 $, by
  % \[ \beta(a, a^\prime) = f(a) a^\prime \quad \text{and} \quad \beta^{-1}(a^\prime) = (1, a^\prime). \]

  % Given a morphism $ s: X \to Y $, we have $ f^*(s): f^*(X) \to f^*(Y) $ given by
  % \[ f^*(s)(x, a^\prime) = (s(x), a^\prime). \]

  % Finally, we have
  % \[ \delta(s) = \lambda x_1 x_2, s(p_1, p_2)(\lambda x_3, x_3 x_1 x_2) \]
  % with $ p_i = \lambda x_1, x_1 (\lambda x_2 x_3, x_{i + 1}) $.

  % For some $ s: U_A^{n + 1} \to U_A $, the left hand side sends $ (a, (a^\prime_1, \dots, a^\prime_n)): f^*(U_A^n) $ to
  % \begin{align*}
  %   &\lambda x_1 x_2, f(s((a_i p_1)_i, p_2)) (\lambda x_3, x_3 x_1 x_2) \circ a^\prime\\
  %   &= \lambda x_1 x_2, f(s((a_i p_1)_i, p_2)) (\lambda x_3, x_3 a^\prime (x_1) x_2)
  % \end{align*}
  % and the right hand side sends it to
  % \begin{align*}
  %   \lambda x_1 x_2, \lambda x_3, (f(s((a_i p_1)_i, p_2))((\lambda x_4, x_4 (a^\prime p_1 (\lambda x_3, x_3 x_1 x_2)) (p_2 (\lambda x_3, x_3 x_1 x_2)))(x_3)))
  % \end{align*}
\end{remark}

\begin{lemma}
  For $ F = \id A $, we have $ \bar F = \id{E(U_A)} $.
\end{lemma}
\begin{proof}
  By the reasoning set forth in Lemma \ref{lem:elementary-faithful}, we only need to check $ \bar F_0 $. Given $ s : \RAct{A_1}(I, U_A) $, we have
  \[ \bar F_0(s) = \gamma^{-1} \cdot F_*(s) \cdot \beta: \RAct{A_1}(I, U_A) \]
  for $ \gamma: F_*(I) \cong I $ and $ \beta: F_*(U_A) \cong U_A $. Then
  \[
    \bar{\id A}(s)(\star) = s(\star) \circ ((\lambda x_1 x_2, x_2) \bullet ()) = s(\star \circ ((\lambda x_1 x_2, x_2) \bullet ())),
  \]
  so $ \bar{\id A}_0(s) = s $ and we can conclude that $ \bar{\id A} = \id{E(U_A)} $.
\end{proof}

\begin{lemma}
  For $ F : \Alg{\Lambda}(A, B) $ and $ G: \Alg{\Lambda}(B, C) $, we have $ \bar F \cdot \bar G = \overline{F \cdot G} $.
\end{lemma}
\begin{proof}
  Again, we only have to check at the terms without free variables. Given $ s : \RAct{A_1}(I, U_A) $, we have
  \[ (\bar F \cdot \bar G)(s) = \gamma_G^{-1} \cdot G_*(\gamma_F^{-1} \cdot F_*(s) \cdot \beta_F) \cdot \beta_G: \RAct{C_1}(I, U_C) \]
  for $ \gamma: F_*(I) \cong I $ and $ \beta: F_*(U_A) \cong U_A $. Then
  \begin{align*}
    (\bar F \cdot \bar G)(s)(\star)
    &= G(F(s(\star)) \circ ((\lambda x_1 x_2, x_2) \bullet_B ())) \circ ((\lambda x_1 x_2, x_2) \bullet_C ())\\
    &= G(F(s(\star \circ ((\lambda x_1 x_2, x_2) \bullet_A ()) \circ ((\lambda x_1 x_2, x_2) \bullet_A ()))))\\
    &= G(F(s(\star)))\\
    &= (F \cdot G)^{-1}(s(\star \circ ((\lambda x_1 x_2, x_2) \bullet_A ())))\\
    &= \overline(F \cdot G)(s)(\star)
  \end{align*}
  so $ (\bar F \cdot \bar G)_0(s) = (\overline{F \cdot G})_0(s) $ and we can conclude that $ \bar F \cdot \bar G = \overline{F \cdot G} $.
\end{proof}

\begin{definition}
  We get a functor from $ \Alg{\Lambda} $ to $ \LamTh $, sending objects $ A $ to $ E_{\RAct{A_1}}(U_A) $ and morphisms $ F: \Alg{\Lambda}(A, B) $ to $ \bar F: E(U_A) \to E(U_B) $.
\end{definition}

\begin{remark}
  Note that for $ X : \R_A $, we have a natural isomorphism of sets with a right $ B_1 $-action
  \[ \psi: \{ f : B \mid F(X) \circ f = f \} \xrightarrow \sim \{ f : A \mid X \circ f = f \} \times B_1 / \sim \]
  given by
  \[ \psi(f) = (X, f) \quad \text{and} \quad \psi^{-1}(f, b) = F(f) \circ b. \]
  % which is $ B_1 $-equivariant and natural in $ X $.
  % We have
  % \[ \psi^{-1}(\psi(f)) = \psi^{-1}(X, f) = F(X) \circ f = f \]
  % \[ \psi(\psi^{-1}(f, b)) = \psi(F(f) \circ b) = (X, F(f) \circ b) = (X \circ f, b) = (f, b) \]
  % We need to show that this is natural in $ X $. Therefore, take $ g : \R_A(X, Y) $. Then $ \R_B(\mathbf 1_1, F(g)) $ is given by postcomposition with $ F(g) $. Also, $ F_*(\R_A(\mathbf 1_1, g)) $ is given by postcomposing with $ g $ on the component $ \R_A(\mathbf 1_1, g) $. Then
  % \[ \psi(F(g) \circ f) = (Y, F(g) \circ f) = (Y \circ g, f) = (g \circ X, f) \]

  Therefore, the following diagram $ 2 $-commutes:
  \begin{center}
    \begin{tikzcd}
      \R_A \arrow[r, "F"] \arrow[d, "{\R_A(\mathbf 1_1, -)}"] & \R_B \arrow[d, "{\R_B(\mathbf 1_1, -)}"]\\
      \RAct{A_1} \arrow[r, "F_*"] & \RAct{B_1}
    \end{tikzcd}
  \end{center}
  Since $ F $ preserves all the structure of $ \R_A $, as do the embeddings of $ \R_A $ and $ \R_B $ into $ \RAct{A_1} $ and $ \RAct{B_1} $, one would expect $ F_* $ to preserves the structure of the full subcategory $ \R_A $ of $ \RAct{A_1} $, including $ U_A $, its finite products and their exponentials. This is another way to argue that $ \bar F $ is indeed a morphism of $ \lambda $-theories.
\end{remark}

\subsection{The unit}

Definining the unit of the adjunction boils down to a version of Scott's representation theorem.

\begin{lemma}
  For a $ \lambda $-theory $ L $, we can define a $ \lambda $-theory isomorphism
  \[ \eta_L: L \xrightarrow{\sim} E(U_{L_0}). \]
\end{lemma}
\begin{proof}
  Recall from Remark \ref{rem:ract-presheaf-equivalence} that we have a chain of equivalences
  \[ \Pshf L \xrightarrow \sim P \mathbf L \xrightarrow \sim P C_{L_1} \xrightarrow \sim P C_{{L_0}_1} \xrightarrow \sim \RAct{{L_0}_1}. \]
  It sends the presheaf $ L^n $ to the set $ L_1^n $ with a right action sending $ s: L_1^n $ and $ t: (L_0)_1 $ to $ (s_i \bullet_L \rho(t))_i : L_1^n $. We have an isomorphism $ s \mapsto (\lambda(s_i))_i: (L_1)^n \to U_{L_0} $, with inverse $ (\rho_i)_i : s \mapsto (\rho(s_i))_i $.

  Then $ \eta_L $ arises by combining this with the isomorphism from Scott's representation theorem (Theorem \ref{thm:representation-theorem})
  \[ \eta_L: L \xrightarrow \sim E_{\Pshf L}(L) \xrightarrow \sim E_{\RAct{{L_0}_1}}(U_{L_0}). \]
  It is quite easy to work out that we can make this explicit as (dropping the $ L $ and just writing $ \eta_L $)
  \[ \eta_n:
    L_n
      \xrightarrow{f \mapsto ((s : \Pshf{L_m^n}) \mapsto f \bullet s)_m}
    \Pshf L(L^n, L)
      \xrightarrow{f \mapsto f_1}
    \RAct{(L_0)_1}(L_1^n, L_1),
      \xrightarrow{\lambda \circ - \circ (\rho_i)_i}
    \RAct{(L_0)_1}(U_{L_0}^n, U_{L_0}),
  \]
  so $ \eta_n(f)(s) = \lambda(f \bullet (\rho(s_i))_i) $ for $ f: L_n $ and $ s: U_{L_0}^n $.
\end{proof}

\begin{remark}
  Note that we can do the same with Scott's version of his representation theorem, using the chain of equivalences and an embedding
  \[ \mathbf R \xrightarrow \sim \mathbf R_{L_0} \xrightarrow{Y} P \R_{L_0} \xrightarrow \sim P C_{{L_0}_1} \xrightarrow \sim \RAct{{L_0}_1}, \]
  Here, it is a bit harder to get an explicit formula for $ \eta $, because we need to do a bit more conversion between $ U_A^n $ and the image of $ U^n $ after the embedding $ \varphi: \mathbb R \hookrightarrow \RAct{{L_0}_1} $. If we quickly define $ \lambda $-terms
  \[ \langle x_i \rangle_i = \langle x_1, \dots, x_n \rangle = \langle \langle \dots \langle (\lambda x_{n + 1} x_{n + 2}, x_{n + 2}), x_1 \rangle, \dots\rangle, x_n \rangle \]
  and corresponding tuples $ (x_i)_i $ and projections $ \pi_i $, we have
  \[ L_n \xrightarrow{f \mapsto \lambda(f \bullet (\pi_i x_1)_i)} \mathbb R(U^n, U) \xrightarrow{f \mapsto f \circ -} \RAct{{L_0}_1}(\varphi(U^n), U_{L_0}) \xrightarrow{(s \mapsto \langle s_i \rangle_i) \circ -} \RAct{{L_0}_1}(U_{L_0}^n, U_{L_0}). \]
  and in the end, we obtain the very same explicit formula
  \[ \eta_n(f)(s) = \iota_{0, n}(f)(x_i)_i \circ (\langle s_i \rangle_i) = \lambda(f \bullet (\rho(s_i))_i). \]
\end{remark}

\begin{lemma}
  $ \eta $ is natural in $ L $. That is, for all $ F: \LamTh(L, L^\prime) $, the following diagram commutes:
  \begin{center}
    \begin{tikzcd}
      L \arrow[r, "F"] \arrow[d, "\eta_L"] & L^\prime \arrow[d, "\eta_{L^\prime}"]\\
      E_{\RAct{{L_0}_1}}(U_{L_0}) \arrow[r, "\bar{F_0}"] & E_{\RAct{{L^\prime_0}_1}}(U_{L^\prime_0})
    \end{tikzcd}
  \end{center}
\end{lemma}
\begin{proof}
  We must show
  \[ \eta_L \cdot \bar{F_0} = F \cdot \eta_{L^\prime}. \]
  Note that for all $ s : L_0 $,
  \begin{align*}
    \bar{F_0}_0(\eta_L(s))
    &= \gamma_{F_0}^{-1} \cdot {F_0}_*(\star \mapsto (\lambda x_1, \iota_{0, 1}(s))) \cdot \beta_F\\
    &= \star \mapsto F_0(\lambda x_1, \iota_{0, 1}(s)) \circ ((\lambda x_1 x_2, x_2) \bullet ())\\
    &= \star \mapsto (\lambda x_1, \iota_{0, 1}(F_0(s)))\\
    &= \eta_{L^\prime}(F_0(s))
  \end{align*}
  and by Lemma \ref{lem:elementary-full}, this concludes the proof.
\end{proof}

\subsection{The counit}

\begin{definition}
  For $ A : \Alg \Lambda $, we define a bijection $ \epsilon_A : E_{\RAct{A_1}}(U_A)_0 \cong A $ as
  \[ \epsilon_A(s) = x_1 (\lambda x_2, x_2) \bullet (s(\star)) \quad \text{and} \quad \epsilon_A^{-1}(a)(\star) = (\lambda x_2, x_1) \bullet a. \]
  These are inverses because $ s(\star) $ is $ A_1 $-equivariant (Lemma \ref{lem:global-action-elements}), and then
  \[ (\lambda x_2, x_1 (\lambda x_3, x_3)) \bullet s(\star) = s(\star) \circ ((\lambda x_1 x_2, x_2) \bullet ()) = s(\star). \]
\end{definition}

We want to show that $ \epsilon_A $ is an isomorphism of $ \Lambda $-algebras. We use Lemma \ref{lem:make-is-lambda-algebra-morphism} for this, so we need to show that it preserves the application and the $ \Lambda $-definable constants.

\begin{lemma}
  We have for all $ a, b: E(U_A)_0 $,
  \[ \epsilon_A((x_1 x_2) \bullet (a, b)) = (x_1 x_2) \bullet (\epsilon_A(a), \epsilon_A(b)). \]
\end{lemma}
\begin{proof}
  For $ a, b: E(U_A)_0 $, we have, using at some point the isomorphism $ \delta: \RAct{A_1}(U_A^{U_A}, A_2) $,
  \begin{align*}
    \epsilon_A((x_1 x_2) \bullet (a, b)) &= (x_1 (\lambda x_3, x_3) (x_2 (\lambda x_3, x_3))) \bullet (a(\star), b(\star))\\
    &= (x_1 x_2) \bullet (\epsilon_A(a), \epsilon_A(b))
  \end{align*}
  and this concludes the proof.
\end{proof}

To show that $ \epsilon $ preserves the $ \Lambda $-definable constants, we first need to show two properties of $ \epsilon $ and $ \eta $:
\begin{lemma}\label{lem:counit-natural}
  $ \epsilon $ is natural in $ A $. That is, for all $ F: \Alg{\Lambda}(A, B) $, the following diagram commutes:
  \begin{center}
    \begin{tikzcd}
      E(U_A)_0 \arrow[d, "\epsilon_A"] \arrow[r, "\bar F_0"] & E(U_B)_0 \arrow[d, "\epsilon_B"]\\
      A \arrow[r, "F"] & B
    \end{tikzcd}
  \end{center}
\end{lemma}
\begin{proof}
  The functor $ F_* : \RAct{A_1} \to \RAct{B_1} $ sends $ X : \RAct{A_1} $ to $ X \times B_1 / \sim : \RAct{B_1} $. We have isomorphisms
  \[ \beta: F_*(U_A) \xrightarrow \sim U_B \quad \text{and} \quad \gamma: F_*(I) \xrightarrow \sim I, \]
  with
  \[ \beta(a, b) = F(a) \circ b \quad \text{and} \quad \gamma^{-1}(\star) = (\star, (\lambda x_1 x_2, x_2) \bullet ()). \]
  Then $ \bar F: \RAct{A_1}(I, U_A) \to \RAct{B_1}(I, U_B) $ is given by
  \[
    \bar F(s)(\star)
    = \gamma^{-1} \cdot (s \times \id{B_1}) \cdot \beta
    = (\lambda x_2, x_1 (\lambda x_3, x_3)) \bullet F(s(\star)),
  \]
  so
  \[
    (\bar F \cdot \epsilon_B)(s)
    = (x_1 (\lambda x_2, x_2)) \bullet F(s(\star))
    = (\epsilon_A \cdot F)(s),
  \]
  which concludes the proof.
\end{proof}

\begin{lemma}\label{lem:zigzag}
  $ \epsilon $ and $ \eta $ satisfy one of the zigzag identities.
\end{lemma}
\begin{proof}
  In this case, the zigzag identity on $ L \mapsto L_0 $ boils down to the following diagram commuting for all $ L : \LamTh $:
  \begin{center}
    \begin{tikzcd}
      L_0 \arrow[rr, "\id{L_0}"] \arrow[rd, "{\eta_L}_0"] & & L\\
      & E(U_{L_0})_0 \arrow[ru, "\epsilon_{L_0}"]
    \end{tikzcd}
  \end{center}
  Now, note that for all $ f : L_0 $,
  \[
    ({\eta_L}_0 \cdot \epsilon_{L_0})(f)
    = (x_1 (\lambda x_2, x_2)) \bullet (\lambda x_1, \iota_{0, 1}(f))
    = f,
  \]
  which shows that the diagram commutes.
\end{proof}

Now, finally, we are ready to show that $ \epsilon $ is an isomorphism of $ \Lambda $-algebras:
\begin{lemma}
  We have for all $ s: \Lambda_0 $,
  \[ \epsilon_A(s \bullet ()) = s \bullet (). \]
\end{lemma}
\begin{proof}
  Consider the following diagram, with $ F: \Alg \Lambda(\Lambda_0, A) $ given by $ F(s) = s \bullet () $:
  \begin{center}
    \begin{tikzcd}
      E(U_{\Lambda_0})_0
        \arrow[r, "\bar F_0"]
        \arrow[d, "\epsilon_{\Lambda_0}"]
      & E(U_A)_0
        \arrow[d, "\epsilon_A"]\\
      \Lambda_0
        \arrow[r, "F"]
        \arrow[u, bend left, "\eta_0"]
      & A
    \end{tikzcd}
  \end{center}
  By Lemma \ref{lem:counit-natural}, the square commutes and by Lemma \ref{lem:zigzag}, we have $ \eta_0 \cdot \epsilon_{\Lambda_0} = \id{\Lambda_0} $.

  Recall that there exists a unique morphism $ \iota_\Lambda: \LamTh(\Lambda, E(U_A)) $, and that for all $ s: \Lambda_0 $, by definition $ s \bullet_{E(U_A)_0} () = \iota_\Lambda(s) \bullet_{E(U_A)} () = \iota_\Lambda(s) $. Since we have $ \eta \cdot \bar F : \LamTh(\Lambda, E(U_A)) $, we must have $ \iota_\Lambda = \eta \cdot \bar F $ and
  \begin{align*}
    \epsilon_A(s \bullet_{E(U_A)} ())
    &= \epsilon_A(\bar{F_0}(\eta_0(s)))\\
    &= F(\epsilon_{\Lambda_0}(\eta_0(s)))\\
    &= F(s)\\
    &= s \bullet_A (),
  \end{align*}
  which concludes the proof.
\end{proof}

\subsection{The equivalence}

By Lemma 3.2 in \autocite{nlab:adjoint_equivalence} and Lemma \ref{lem:zigzag}, $ \eta $ and $ \epsilon $ satisfy both zigazg identities, and we can state the fundamental theorem of the $ \lambda $-calculus:
\begin{theorem}
  There is an adjoint equivalence $ \LamTh \cong \Alg{\Lambda} $, sending a $ \lambda $-theory $ L $ to the $ \Lambda $-algebra $ L_0 $, with an inverse functor that sends a $ \Lambda $-algebra $ A $ to the theory $ E_{\RAct{A_1}}(U_A) $.
\end{theorem}

\begin{remark}
  Hyland remarks that the isomorphism $ U_A^{U_A} \cong A_2 $ can be generalized to isomorphisms $ U_A^{U_A^n} \cong A_{n + 1} $.
  \begin{align*}
    \RAct{A_1}(U_A^n, U_A)
    &\cong \RAct{A_1}(I, U_A^{U_A^n})\\
    &\cong \RAct{A_1}(I, A_{n + 1})\\
    &\cong \{ f : A_{n + 1} \mid \forall a : A_1, f \circ a = f \}\\
    &\cong A_n.
  \end{align*}
  % In particular, we can pick $ a = \lambda x_1, x_1 (\lambda x_2, x_2) $. Then we have an isomorphism from this subset of $ A_{n + 1} $ to $ A_n $, given by
  % \[ f \mapsto (x_1 (\lambda x_2, x_2)) \bullet f \quad \text{and} \quad g \mapsto (\lambda x_2, x_1) \bullet g. \]
  Explicitly, we get an isomorphism $ \psi: \RAct{A_1}(U_A^n, U_A) \xrightarrow \sim A_n $, given by
  \[ \psi(f) = (\lambda x_2 \dots x_{n + 1}, x_1 (x_2, \dots, x_{n + 1})) \bullet (f (\pi_1, \dots, \pi_n)) \]
  and
  \[ \psi^{-1}(g)(a_1, \dots, a_n) = (\lambda x_{n + 2}, x_1 (x_2 x_{n + 2}) \dots (x_{n + 1} x_{n + 2})) \bullet (g, a_1, \dots, a_n). \]
  Note that to show this, we need to use the $ A_1 $-equivariance of $ f $ at some point:
  \begin{align*}
    &f(\pi_2, \dots, \pi_{n + 1})((\lambda x_{n + 1}, x_{n + 1}), x_1, \dots, x_n)\\
    &= (f(\pi_1, \dots, \pi_n) \circ \langle \pi_2, \dots, \pi_{n + 1} \rangle) ((\lambda x_{n + 2}, x_{n + 2}), x_1, \dots, x_n)\\
    &= f(\pi_1, \dots, \pi_n) (x_1, \dots, x_n).\\
  \end{align*}
  This gives a $ \lambda $-theory structure on $ (A_n)_n $, with
  \begin{align*}
    y_{n, i} &= (\lambda x_1 \dots x_n, x_{n, i}) \bullet ();\\
    f \bullet g &= (\lambda x_{m + 2} \dots x_{m + n + 1}, x_1 (x_2 x_{m + 2} \dots x_{m + n + 1}) \dots (x_{m + 1} x_{m + 2} \dots x_{m + n + 1})) \bullet (f, g_1, \dots, g_m);\\
    \rho(f) &= \mathbf 1_m \circ f;\\
    \lambda(h) &= f.
  \end{align*}
  for $ f: A_m $, $ g: A_n^m $ and $ h: A_{m + 1} $.

  Then any $ F: \Alg \Lambda(A, B) $ gives functions $ F: A_n \to B_n $, and these give a $ \lambda $-theory morphism in $ \LamTh((A_n)_n, (B_n)_n) $.

  The natural isomorphism $ \epsilon_A: A_0 \to A $ is just $ \id A $. Note that, even though $ A_0 = A $ as sets, their $ \Lambda $-algebra structures are defined differently, so it takes some work to show that $ \epsilon_A $ is a $ \Lambda $-algebra morphism. Also, $ {\eta_L}_n : L_n \to (L_0)_n $ is given by $ \lambda^n $, with $ \rho^n $ as its inverse. The zigzag identities are trivial, so this gives another, very elemental proof of the fundamental theorem.
  % \begin{align*}
  %   \psi^{-1}(\psi(f))(b)
  %   &= (\lambda x_{n + 2}, x_1 (x_2 x_{n + 2}) \dots (x_{n + 1} x_{n + 2})) \bullet (\psi(f), b_1, \dots, b_n)\\
  %   &= (\lambda x_{n + 2}, x_1 (x_2 x_{n + 2}, \dots, x_{n + 1} x_{n + 2})) \bullet (f (\pi_1, \dots, \pi_n), b_1, \dots, b_n)\\
  %   &= (x_1 \circ \langle x_2, \dots, x_{n + 1} \rangle) \bullet (f (\pi_1, \dots, \pi_n), b_1, \dots, b_n)\\
  %   &= f (\pi_1, \dots, \pi_n) \circ \langle b_1, \dots, b_n \rangle\\
  %   &= f (\pi_1 \circ \langle b_1, \dots, b_n \rangle, \dots, \pi_n \circ \langle b_1, \dots, b_n \rangle)\\
  %   &= f (b_1, \dots, b_n).
  % \end{align*}
  % \begin{align*}
  %   \psi(\psi^{-1}(g))
  %   &= (\lambda x_2 \dots x_{n + 1}, x_1 (x_2, \dots, x_{n + 1})) \bullet ((\lambda x_{n + 2}, x_1 (x_2 x_{n + 2}) \dots (x_{n + 1} x_{n + 2})) \bullet (g, \pi_1, \dots, \pi_n))\\
  %   &= (\lambda x_2 \dots x_{n + 1}, x_1 x_2 \dots x_{n + 1}) \bullet g\\
  %   &= g.
  % \end{align*}
\end{remark}

\section{Theory of extensions}

The fundamental theorem of the $ \lambda $-calculus that Hyland shows is actually not of the form shown above. To get there, we first need to show that the category of $ T $-algebras for an algebraic theory $ T $ has coproducts, and define the `theory of extensions'.

Let $ \llbracket n \rrbracket $ denote the finite set $ \{ 1, 2, \dots, n \} $. Let \index{$ \F $}$ \F \subseteq \FinSET $ be the full subcategory with $ \F_0 = \{ \llbracket 0 \rrbracket, \llbracket 1 \rrbracket, \dots \} $.

\begin{lemma}
  Let $ T $ be an algebraic theory. The category of $ T $-algebras has coproducts.
\end{lemma}
\begin{proof}
  This is shown in \autocite{algebaric-theories-2010}, in the lemmas leading up to Theorem 4.5.

  Explicitly, we can express the coproduct of algebras, and especially its set, as the following coend (\autocite{Hyland}, Proposition 2.5) (see also Section \ref{sec:coends} for more on coends)
  \[ A + B = \int^{(m, n) : \F \times \F} A^m \times B^n \times T_{m^\prime + n^\prime}, \]
  considering $ T $ as a functor $ T : \F \to \SET $ as a functor (see Remark \ref{rem:abstract-clone-functor}). Note that we do not need the exact definition of $ A + B $ for the rest of this section. Nonetheless, it is interesting to see how it is defined and why this definition works.

  One can think of $ A + B $ consisting of elements $ t \bullet (a + b) $ for $ t: T_{m + n} $, $ a : A^m $ and $ b: B^n $ (writing $ (a + b) $ for $ (a_1, \dots, a_m, b_1, \dots, b_n) $), `substituting' the $ a_i $ and $ b_j $ for the $ x_i $ and $ x_{m + j} $ in $ t $.

  The relations then give `associativity' of this substitution and also assure that reordering or duplicating $ x_i $ and their corresponding $ a_i $ and $ b_j $ do not yield different elements: associating the images on the left and right of
  \[ A^{m^\prime} \times B^{n^\prime} \times T_{m^\prime + n^\prime} \leftarrow A^{m^\prime} \times B^{n^\prime} \times T_{m + n} \rightarrow A^m \times B^n \times T_{m + n} \]
  gives
  \[ (t \bullet ((x_{f(i)})_i + (x_{m + g(j)})_j)) \bullet a + b = t \bullet ((a_{f(i)})_i + (b_{g(j)})_j) \]
  for $ f: \llbracket m \rrbracket \to \llbracket m^\prime \rrbracket $, $ g: \llbracket n \rrbracket \to \llbracket n^\prime \rrbracket $, $ t : T_{m + n} $, $ a : A^{m^\prime} $ and $ b : B^{n^\prime} $.

  Then it becomes clear what the action of $ T $ will be on $ A + B $, although the precise definition has quite a combinatorial nature. For $ s : T_l $, $ t_i : T_{m_i + n_i} $, $ a_i : A^{m_i} $ and $ b_i : B^{n_i} $, define the disjoint embeddings
  \[ d_i : \left\llbracket m_i + n_i \right\rrbracket \cong \left\llbracket m_i \right\rrbracket \sqcup \left\llbracket n_i \right\rrbracket \hookrightarrow \left\llbracket \sum_j m_j \right\rrbracket \sqcup \left\llbracket \sum_j n_j \right\rrbracket \cong \left\llbracket \sum_j m_j + \sum_j n_j \right\rrbracket, \]
  which we will use to make sure that the $ x_j $ in the different $ t_i $ are mapped to distinct variables. Then we can define
  \[ s \bullet (t_i \bullet a_i + b_i) = (s \bullet (t_i \bullet (x_{d_i(j)})_j)_i) \bullet (a_1 + b_1 + \dots + a_l + b_l). \]
  More formally (using the coend injections $ A^{\sum_k m_k} \times B^{\sum_k n_k} \times T_{\sum_k m_k + \sum_k n_k} \to A + B $), this gives functions
  \[ T_l \to (A^{m_1} \times B^{n_1} \times T_{m_1 + n_1} \to (\dots \to (A^{m_l} \times B^{n_l} \times T_{m_l + n_l} \to A + B) \dots )), \]
  commuting with the relations between the different $ (A^m \times B^n \times T_{m + n}) $, which, by repeatedly using the universal property of the coend, then correspond to functions
  \[ T_l \to (A + B \to (\dots \to (A + B \to A + B) \dots )), \]
  or, equivalently, a function
  \[ T_l \times (A + B)^l \to A + B. \]

  We have left and right injections $ A \to A + B $ and $ B \to A + B $, given respectively by the maps $ A^1 \times B^0 \times T_{1 + 0} \to A + B $ and $ A^0 \times B^1 \times T_{0 + 1} \to A + B $:
  \[ a \mapsto x_1 \bullet a \quad \text{and} \quad b \mapsto x_1 \bullet b \]
  and every element $ A + B $ arises by the action of $ t $ on combinations of these embedded elements:
  \[ t \bullet (a + b) = t \bullet ((x_1 \bullet a_i)_i + (x_1 \bullet b_j)_j), \]
  which ultimately can be used to show that $ A + B $ indeed has the universal property of the coproduct.
\end{proof}

\begin{definition}[Theory of extensions]
  Let $ T $ be an algebraic theory and $ A $ a $ T $-algebra. We can define an algebraic theory $ T_A $ called `the theory of extensions of $ A $' with $ (T_A)_n = T_n + A $. The left injection of the variables $ x_i : T_n $ gives the variables.
  Now, take $ h: (T_n + A)^m $. Sending $ g: T_m $ to $ g \bullet h $ gives a $ T $-algebra morphism $ T_m \to T_n + A $. Together with the right injection morphism of $ A $ into $ T_n + A $, this gives us a $ T $-algebra morphism from the coproduct: $ T_m + A \to T_n + A $. In particular, we have a function on sets $ (T_m + A) \times (T_n + A)^m \to T_n + A $, which we will define our substitution to be.
\end{definition}