### 2024-09-09 report/src/metadata.tex
Update some notation for the Yoneda lemma

% !TEX root = document.tex

\title{Universal algebra and the \texorpdfstring{$ \lambda $}{lambda}-calculus}
\subtitle{Version of \today}
%\subtitle{Master's Thesis}
\author{Arnoud van der Leer}
\authoremail{\url{a.a.vanderleer@student.tudelft.nl}}
\birthplace{Leiderdorp, the Netherlands}
\studentid{4485890}
\coverpicture{
  \centerline{
    \Large
    \begin{tikzcd}[
      ampersand replacement=\&,
      sep=scriptsize
    ]
      \&
      C_{{L_0}_1} \arrow[dd, "\yo" description, hook] \arrow[rrr, "\iota_{C_{{L_0}_1}}" description, hook] \arrow[rd, "\sim" description, no head] \& \& \& \R_{L_0} \arrow[dd, "\yo" description, hook] \arrow[rd, "\sim" description, no head] \& \& \\
      \&
      \& C_{L_1} \arrow[dd, "\yo" description, hook] \arrow[r, hook] \& \mathbf L \arrow[rr, hook] \arrow[dd, "\yo" description, hook] \& \& \mathbf R \arrow[dd, "\yo" description, hook] \arrow[rdd, hook] \& \\
      \RAct{{L_0}_1} \&
      P C_{{L_0}_1} \arrow[l, "\sim"'] \arrow[rd, "\sim" description, no head] \& \& \& P \R_{L_0} \arrow[lll, "\sim" description] \arrow[rd, "\sim" description, no head] \& \& \\
      \&
      \& P C_{L_1} \& P \mathbf L \arrow[l, "\sim" description] \& \& P \R \arrow[ll, "\sim" description] \& \Pshf L \arrow[lll, "\sim" description, bend left]
    \end{tikzcd}
  }
}
\colophon{\noindent
  \copyright{} \the\year{} \: \theauthor. \\[1em]
  Cover picture: A 2-commutative diagram featuring categories that appear in this thesis.
}