### 2024-11-08 910-alternative-definitions.tex
Add chapter introductions

\chapter{Alternative definitions}\label{ch:alternative-definitions}
In the literature, there are many different but equivalent definitions, carrying many different names, for the objects that are called `algebraic theories' in Section \ref{sec:algebraic-theories}. To make matters even more confusing, the same name can refer to different things in different parts of the literature. This appendix will showcase various definitions and give them a name. Of particular interest is the Lawvere theory (Section \ref{sec:lawvere-theory}), which is also used in Hyland's paper.

In this appendix, we will denote the finite set $ \{ 1, \dots, n \} $ as $ \llbracket n \rrbracket $.

\section{Abstract Clone}
\begin{definition}
  An algebraic theory as presented in Section \ref{sec:algebraic-theories}, is usually called an \iindex{abstract clone}. In this thesis, outside of this specific section, we will call it algebraic theory to be consistent with the names that Hyland attaches to objects.
\end{definition}

\begin{remark}\label{rem:abstract-clone-functor}
  The definition of algebraic theory that Hyland gives is closest to that of an abstract clone. However, instead of a sequence of sets $ (T_n)_n $, he requires a functor $ T: F \to \SET $ (with $ F \subseteq \FinSET $ the skeleton category of finite sets $ F_0 = \{ \llbracket 0 \rrbracket, \llbracket 1 \rrbracket, \dots \} $), and he requires $ \bullet: T_m \times T_n^m \to T_n $ to be dinatural in $ n $ and natural in $ m $.

  Using naturality, one can show that with such a functor we have $ x_{n, i} = f(a)(x_{1, 1}) $ for the function $ a(1) = i : \llbracket n \rrbracket $.

  Alternatively, using the same naturality, one can show that this functor sends a morphism $ a: \llbracket m \rrbracket \to \llbracket n \rrbracket $ to the function $ T_m \to T_n $ given by
  \[ f \mapsto f \bullet (x_{n, a(i)})_i. \]
  If we take this to be the definition of our functor on morphisms, the (di)naturality in $ m $ and $ n $ can be shown using the associativity and the laws about the interaction between $ \bullet $ and the $ x_i $.

  Since any additional properties mean extra complexity when formalizing, and since the proofs rarely use the functor structure, we decided to reduce the functor $ T: \FinSET \to \SET $ to a sequence of sets $ (T_n)_n $.
\end{remark}

\section{Lawvere Theory}\label{sec:lawvere-theory}
\begin{definition}
  An \iindex{algebraic theory} as presented in \autocite{algebraic-theories-2010} is a small category with finite products.
\end{definition}
\begin{definition}
  An \iindex{algebra} for an algebraic theory $ T $ is a finite-product-preserving functor $ T \to \SET $.
\end{definition}

This definition is more general than the definition of algebraic theory in Section \ref{sec:algebraic-theories}. To make it equivalent, we have to be more specific about the objects of the category:
\begin{definition}\label{def:lawvere-theory}
  A \iindex{Lawvere theory}, or \textit{one-sorted algebraic theory} is a category $ L $, with $ L_0 = \{ 0, 1, \dots \} $, such that $ n = 1^n $, the $ n $-fold product.
\end{definition}

\begin{lemma}[One direction: \coqident{AlgebraicTheories.AlgebraicTheoryToLawvereTheory}{algebraic_theory_to_lawvere}]\label{lem:lawvere-clone}
  There is an equivalence between abstract clones and Lawvere theories.
\end{lemma}
\begin{proof}
  Let $ C $ be an abstract clone. We construct a Lawvere theory $ L $ as follows: We have objects $ L_0 = \{ 0, 1, \dots \} $ and morphisms $ L(m, n) = C_m^n $. The identity morphism is $ \id n = (x_i)_i : L(n, n) $ and for $ f: L(l, m) $, $ g: L(m, n) $, we have composition
  \[ f \cdot g = (g_i \bullet f)_i: L(l, n). \]
  Lastly, we have product projections $ \pi_{n, i} = x_{n, i}: L(n, 1) $ for all $ 1 \leq i \leq n $.

  Conversely, if $ L $ is a Lawvere theory, we construct an abstract clone $ C $ as follows:
  We take $ C_n = L(n, 1) $. We take the $ x_{n, i} $ to be the product projections $ \pi_i : L(n, 1) $ and we define the substitution $ f \bullet g = \langle g_i \rangle_i \cdot f $ the composite of the product morphism $ \langle g_i \rangle_i $ with $ f $.
\end{proof}

\subsection{Algebras for Lawvere Theories}

\begin{lemma}\label{lem:lawvere-algebra}
  Let $ C $ be an abstract clone, and $ L $ be its associated Lawvere theory by the equivalence given above. A $ C $-algebra is equivalent to an algebra for $ L $.
\end{lemma}
\begin{proof}
  Let $ A $ be a $ C $-algebra. We will construct a functor $ F: L \to \SET $ as follows: We take $ F(n) = A^n $. We define the action on morphisms as
  \[ F(f)(a) = (f_i \bullet a)_i \]
  for $ f: L(m, n) = L(m, 1)^n $ and $ a: A^m $.

  Conversely, let $ F: L \to \SET $ be a functor. We take the $ C $-algebra $ A $, with $ A = F(1) $ and for $ f: C_n = L(n, 1) $ and $ a: A^n $, $ f \bullet a = F(f)(a) $.
\end{proof}

\subsection{Presheaves for Lawvere Theories}

\begin{lemma}[\coqident{AlgebraicTheories.AlgebraicTheoryToLawvereTheory}{algebraic_presheaf_weq_lawere_presheaf}]\label{lem:lawvere-presheaf}
  Let $ C $ be an abstract clone, and $ L $ be its associated Lawvere theory by the equivalence given above. A $ C $-presheaf is equivalent to a presheaf over $ L $.
\end{lemma}
\begin{proof}
  The correspondence between $ C $-presheaves $ P $ and presheaves $ Q $ over $ L $ is as follows:

  The sets $ P_n $ correspond to the action of $ Q $ on objects $ Q(n) $.

  The $ P $-action $ P_m \times C_n^m \to P_n $ corresponds to the action of $ Q $ on morphisms $ L(n, m) \times Q(m) \to Q(n) $.
\end{proof}

\section{Cartesian Operad}

\begin{definition}
  A \iindex{cartesian operad} T is a functor $ T: F \to \SET $, together with an `identity' element $ \id{}: T(1) $ and for all $ m, n_1, \dots, n_m: \mathbb N $, a composition map
  \[ T(m) \times \prod_i T(n_i) \to T\left(\sum_i n_i\right), \]
  written $ (f, g_1, \dots, g_m) \mapsto f[g_1, \dots, g_m] $, satisfying some identity, associativity and naturality conditions. See \autocite[][Definition 2.1]{lambda-monoid} for more details.
\end{definition}

\begin{remark}
  One can arrive at this concept as a standalone definition, or one can view a cartesian operad as a \iindex{cartesian multicategory} with one element. A multicategory is a category in which morphisms have type $ C((X_1, X_2, \dots, X_n), Y) $ instead of $ C(X, Y) $. A cartesian multicategory is a multicategory in which one can permute the $ X_i $ of a morphism and has `contraction' and `weakening' operations:
  \begin{align*}
    C((X_1, \dots, X_i, X_i, \dots, X_n), Y) &\to C((X_1, \dots, X_i, \dots, X_n), Y),\\
    C((X_1, \dots, X_{i - 1}, X_{i + 1}, \dots, X_n), Y) &\to C((X_1, \dots, X_{i - 1}, X_i, X_{i + 1}, \dots, X_n), Y).
  \end{align*}
\end{remark}

\begin{lemma}
  There is an equivalence between abstract clones and cartesian operads.
\end{lemma}
\begin{proof}
  Let $ C $ be an abstract clone. We define a cartesian operad $ T $ with $ T(n) = C_n $ and $ T(f)(t) = t \bullet (x_{f(1)}, \dots, x_{f(m)}) $ for $ f: \llbracket m \rrbracket \to \llbracket n \rrbracket $ and $ t: C_m $. The identity element is $ x_{1, 1} $ and the composition $ f[g_1, \dots, g_n] $, for $ f: C_m $ and $ g_i: C_{n_i} $, is given by lifting all terms to $ C_{\sum_i n_i} $ and then substituting:
  \[ f[g_1, \dots, g_n] = f \bullet (T(\iota_i)(g_i)) \]
  for $ \iota_i : \llbracket n_i \rrbracket \hookrightarrow \llbracket \sum_i n_i \rrbracket $ the pairwise disjoint injections.

  Conversely, given a cartesian operad $ T $, we construct an abstract clone $ C $ with $ C_n = T(\llbracket n \rrbracket) $. The variables are $ x_{n, i} = \iota_{n, i}(\id{}) $, for $ \iota_{n, i}: \llbracket 1 \rrbracket \hookrightarrow \llbracket n \rrbracket $ the morphism that sends $ 1 $ to $ i $. The substitution $ f \bullet (g_i)_i $ for $ g_1, \dots, g_m: T(n) $ is given by composing and then identifying some variables:
  \[ f \bullet (g_i)_i = T(\pi)(f[g_1, \dots, g_m]) \]
  for $ \pi: \llbracket m n \rrbracket \to \llbracket n \rrbracket $ the function that sends $ i + 1 $ to $ (i \mod n) + 1 $.
\end{proof}

\section{Relative Monad}
\begin{definition}
  Let $ S: C \to D $ be a functor. A \iindex{relative monad} on $ S $ is a functor $ T: C \to D $, together with a natural transformation $ \eta: S \Rightarrow T $ and a `kleisli extension' $ (-)^*: D(S(X), T(Y)) \to D(T(X), T(Y)) $, natural in both $ S $ and $ T $, such that for all $ f: D(S(X), T(Y)) $ and $ g: D(S(Y), T(Z)) $,
  \[ \eta_X^* = \id{T X}, \qquad f = \eta_X \cdot f^* \qquad \text{and} \qquad (f \cdot g^*)^* = f^* \cdot g^*. \]
\end{definition}

\begin{remark}
  Note that for an adjunction $ F \dashv G $, $ F \bullet G $ gives a monad. In the same way, there exists a notion of \iindex{relative adjunction}, from which we can obtain a relative monad \autocite[][Theorem 2.10]{monads-endofunctors}.
\end{remark}

\begin{remark}
  Now, there is a result that states: There is an equivalence between abstract clones and relative monads on the embedding $ \iota: \FinSET \hookrightarrow \SET $.

  Note that the objects of $ \FinSET $ are defined to be sets $ X $, together with a proof that there exists some $ n: \mathbb N $ and some bijection $ f: X \xrightarrow{\sim} \llbracket n \rrbracket $. This existence of $ n $ and $ f $ is given by the propositional truncation $ \left\Vert \sum_{n: \mathbb N}, X \xrightarrow{\sim} \llbracket n \rrbracket \right\Vert $.

  Classically, this construction starts with ``fix, for all $ X : \FinSET $, a bijection $ f: X \xrightarrow{\sim} \llbracket n \rrbracket $''. However, since $ X $ only provides \textit{mere existence} of such a bijection without choosing one (by the propositional truncation), there is no way to obtain $ f $ without using the axiom of choice.

  Now, we can partially circumvent this problem by noting that we have a fully faithful and essentially surjective embedding $ F \to \FinSET $ (for $ F $ a skeleton of finite sets), which induces an adjoint equivalence $ [F, \SET] \xrightarrow{\sim} [\FinSET, \SET] $, so we can lift the functor part of a relative monad on $ F \to \SET $ to $ \FinSET \to \SET $. We can also lift the natural transformation using this equivalence. However, the kleisli extension cannot be lifted using this and we are stuck.

  Therefore, we will prove a modified statement:
\end{remark}

\begin{lemma}
  There is an equivalence between abstract clones and relative monads on the embedding $ \iota: F \hookrightarrow \SET $.
\end{lemma}
\begin{proof}
  Let $ C $ be an abstract clone. We define a relative monad $ T $ as follows: We take $ T(\llbracket n \rrbracket) = C_n $. For a morphism $ a: F(\llbracket m \rrbracket, \llbracket n \rrbracket) $. We take $ T(a)(f) = f \bullet (x_{a(i)})_i $. We define $ \eta_{\llbracket n \rrbracket}(i) = x_{n, i} $. Finally, for $ g: \SET(\llbracket m \rrbracket, T(\llbracket n \rrbracket)) $, we define $ g^*(f) = f \bullet g $.

  Conversely, let $ (T, \eta, (\cdot)^*) $ be a relative monad on the embedding $ \iota: F \hookrightarrow \SET $. We define an abstract clone $ C $ with $ C_n = T(\llbracket n \rrbracket) $. Substitution is defined as $ f \bullet g = g^*(f) $ for $ f: C_m $ and $ g: C_n^m $ and we have variables $ x_{n, i} = \eta_{\llbracket n \rrbracket}(i) $.
\end{proof}

\section{Monoid in a Skew-Monoidal Category}
For details and a general treatment, see \autocite{monads-endofunctors}, Section 3 and specifically Theorem 3.4.

\begin{definition}
  Consider the category $ [F, \SET] $. Note that we have a functor $ \iota: F \hookrightarrow \SET $ and that $ \SET $ has colimits. Therefore, we can define a `tensor product' on functors $ [F, \SET] $ as
  \[ F \otimes G = G \bullet \Lan{\iota}{F}. \]
  Together with the `unit' $ \iota $, this gives $ [F, \SET] $ a `skew-monoidal category' structure.
\end{definition}

\begin{definition}
  Given a (skew-)monoidal category $ (C, \otimes, I) $, a \iindex{monoid} in this category is an object $ T : C $, together with a `multiplication' $ \mu: C(T \otimes T, T) $ and a `unit' morphism $ \eta: C(I, T) $ satisfying a couple of laws \autocite[][Section III.6]{MacLane}.
\end{definition}

\begin{lemma}
  There exists an equivalence between relative monads on $ \iota: F \hookrightarrow \SET $ and monoids in the skew-monoidal category $ [F, \SET] $.
\end{lemma}
\begin{proof}
  A monoid in $ [F, \SET] $ consists of an object $ T: [F, \SET] $, together with natural transformations $ \mu: T \otimes T \Rightarrow T $ and $ \eta: \iota \Longrightarrow T $. We can immediately see the functor $ T $ and natural transformation $ \eta $ of the relative monad pop up here. The kleisli extension corresponds to $ \mu $; they are related to each other via the properties of the left Kan extension of $ T $.
\end{proof}