### 2024-11-26 000-abstract.tex
Write an abstract

The $ \lambda $-calculus is a versatile tool both in mathematical logic and computer science. This thesis studies and expands upon Martin Hyland's paper `Classical lambda calculus in modern dress'. It gives examples for the definitions and provides more detailed proofs, as well as one new proof for Hyland's fundamental theorem of the $ \lambda $-calculus. It complements these definitions and proofs with material of previous authors by which Hyland has been inspired.
The thesis translates Hyland's paper from set theory with classical logic to univalent foundations, and showcases where subtleties arise in such a translation. In particular, it discusses the different implementations of the Karoubi envelope in univalent foundations.
Lastly, it discusses the accompanying formalization of parts of Hyland's paper, with in particular a tactic that was developed for applying $ \beta $-reduction and substitution to $ \lambda $-terms.